\documentclass{article}
\input{theorem}
\input{prooftree}
\usepackage{stmaryrd}
\input{linear}
\def\erule#1#2{\begin{prooftree}#1\justifies #2\end{prooftree}}
\def\lpar{\bindnasrepma}
\def\lamp{\binampersand}
\def\bpush{\mathbf{push}}
\newarrow {Equiv} =====
\def\FinSet{\mathbf{FinSet}}
\def\x{\times}
\def\o{\comp}
\def\ups#1{#1^G}
\def\dns#1{#1^B}
\def\rsn{\ \mathsf{sn}}
\def\rx{\mapsto}
\def\ep{\varepsilon}
\def\wat{{@}}
\def\sem#1{[\![#1]\!]}
\def\cn{{:}}
\def\rok{\mathrel\mathsf{ok}}
\def\rtype{\mathrm{type}}
\def\rset{\mathbf{Set}}
\def\rcat{\mathbf{Cat}}
\def\rid{\mathrm{id}}
\def\C{\mathbf{C}}
\def\R{\mathbf{R}}
\def\E{\mathbf{E}}
\def\T{\mathbf{T}}
\def\B{\mathbf{B}}
\def\F{\mathbb{F}}
\def\H{\mathbf{H}}
\def\pair#1#2{\langle#1,#2\rangle}
\def\tri{\triangleright}
\def\o{\circ}
\def\rctx{\,\mathrm{ctx}}
\def\del{\partial}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\def\also#1{\ \textcolor{blue}{\celse #1}}
\usepackage{tikz}
\usepackage{tikz-cd}
\newcounter{nodemaker}
\setcounter{nodemaker}{0}
\def\twocell#1#2{%
  \global\edef\mynodeone{twocell\arabic{nodemaker}}%
  \stepcounter{nodemaker}%
  \global\edef\mynodetwo{twocell\arabic{nodemaker}}%
  \stepcounter{nodemaker}%
  \ar[#1,phantom,shift left=3,""{name=\mynodeone}]%
  \ar[#1,phantom,shift right=3,""'{name=\mynodetwo}]%
  \ar[Rightarrow,#2,from=\mynodeone,to=\mynodetwo]%
}
\def\twocellswap#1#2{%
  \global\edef\mynodeone{twocell\arabic{nodemaker}}%
  \stepcounter{nodemaker}%
  \global\edef\mynodetwo{twocell\arabic{nodemaker}}%
  \stepcounter{nodemaker}%
  \ar[#1,phantom,shift right=3,""{name=\mynodeone}]%
  \ar[#1,phantom,shift left=3,""'{name=\mynodetwo}]%
  \ar[Rightarrow,#2,from=\mynodeone,to=\mynodetwo]%
}

\def\maps{\ar[dd,mapsto,shorten <=2em, shorten >=2em]}
\author{Jason Reed}
\begin{document}

\tikzset{
   commutative diagrams/.cd,
   arrow style=tikz,
   diagrams={>=stealth}}
% got this from
% https://tex.stackexchange.com/questions/169512/tikz-style-arrow-tips-missing-when-using-tikz-cd-crossing-over
% ???

\section{Definitions}
Let $P(X)$ be the free finitely complete category on $X$, for $X$ a set or
category or whatever.

\subsubsection*{Meshes}
A {\em mesh} $M$ is
\begin{itemize}
\item a finitely complete category $M_0$ (of `objects/types')
\item a finitely complete category $M_1$ (of `morphisms/relations')
\item a bifibration $\del_M : M_1 \to P(M_0)$ (`domain/boundary')
\end{itemize}
such that $\del_M$ preserves finite limits, and satisfies Beck-Chevalley,
and the product of opcartesian morphisms is opcartesian.

We may sometimes when unambiguous treat name of a mesh's total category $M_1$ as the
name of the mesh itself.

\subsubsection*{Mesh Morphisms}
A mesh morphism $M \to N$ is
\begin{itemize}
\item a finite limit preserving functor $F_0 : M_0 \to N_0$
\item a finite limit, cartesian morphism, and opcartesian morphism-preserving
functor $F_1 : M_1 \to N_1$
\end{itemize}
such that
\[
\begin{tikzcd}
M_1 \ar[r, "F_1"] \ar[d, "\del_M" swap]  & N_1\ar[d, "\del_N"] \\
M_0 \ar[r, "F_0" swap] & N_0
\end{tikzcd}
\]

\subsubsection*{Cartesian Meshes}

A c-mesh is a mesh for which the base category is $P(X)$ for some set $X$.
A c-mesh morphism is one for which the functor between base categories
arises as $P(f)$ for some set function $f$.

\subsubsection*{Free C-Mesh on a Set}

The free $c$-mesh on a set $X$ is the codomain bifibration
$\begin{tikzcd}[cramped] P(X)^\to \ar[r, "\cod"]& P(X)\end{tikzcd}$.


\subsubsection*{Free C-Mesh on a Category}

To construct the free c-mesh $M(\C)$ on a category $\C$, take the pullback

\[
\begin{tikzcd}
M(\C) \ar[rd, phantom, very near start, "\lrcorner"] \ar[r, ""] \ar[d, "\del" swap]  & P(\C)^\to\ar[d, "\cod"] \\
P(\C_0) \ar[r, "\iota" swap, hook] & P(\C)
\end{tikzcd}
\]


where $\iota$ is the evident inclusion.

\subsubsection*{Free C-Mesh on a single $n$-ary Relation}

Let $\R_n$ be the free finitely complete category with $n+1$ objects $A_0, \ldots, A_{n-1}, R$ and one morphism
$\theta : R \to A_0 \times \cdots \times A_{n-1}$.
The free c-mesh $\sharp n$ on a single $n$-ary relation is given by the pullback
\[
\begin{tikzcd}
\sharp n \ar[rd, phantom, very near start, "\lrcorner"] \ar[r, ""] \ar[d, "\del" swap]  & \R_n^\to\ar[d, "\cod"] \\
P(n) \ar[r, "\iota" swap, hook] & \R_n
\end{tikzcd}
\]
where $\iota$ is the evident inclusion.



Suppose I have a opfibration $f_1 : a_1 \to b_1, f_2 : a_2 \to b_2$ over $F_1, G_1 : A \to B$.
\section{Lemmas}
\begin{lemma}
Every opfibration embeds into a codomain-ish opfibration such that
the embedding is a pullback square.
\end{lemma}

\begin{proof}
Let $p : \E \to \B$ be given. Let's construct the category $\E:\B$.
Its objects are the objects of $\B$ plus the objects of $\E$.
Its morphisms $E \to E$
Its morphisms $E \to B$ are all the $e$-morphisms $E \to E'$ such that $pE' = B$.
Its morphisms $B \to B$ are identities.



%% It has for morphisms from $[e_1, E_1, b_1]: E'_1 \to B_1$ to
%% $[e_2, E_2, b_2]: E'_2 \to B_2$ the triples $(e', e, b)$ such that
%% \[
%% \begin{tikzcd}
%% E_1' \ar[r, "e'"] \ar[d, "e_1"] &E_2' \ar[d, "e_2"]\\
%% E_1 \ar[r, "e"]& E_2 \\
%% \end{tikzcd}
%% \qquad
%% \begin{tikzcd}
%% pE_1 \ar[r, "pe"] \ar[d, "b_1"] &pE_2 \ar[d, "b_2"]\\
%% B_1 \ar[r, "b"]& B_2 \\
%% \end{tikzcd}
%% \]
%% commute.
Then we claim
\[
\begin{tikzcd}
E \ar[r] \ar[d] &B' \ar[d]\\
B \ar[r, equal]& B \\
\end{tikzcd}
\qquad
\begin{tikzcd}
B_1 \ar[r] \ar[d] &B_2 \ar[d]\\
B \ar[r, equal]& B \\
\end{tikzcd}
\]
An object of $(\E : \B)^\Delta$ is a pair of $B \in \B$ and $f : (E : B)^\to$
such that $\cod f = B$; i.e. a diagonal morphism $E \to B$ or a $B$-morphism $B_0 \to B$.
\[
\begin{tikzcd}
\E \ar[d, "p" swap] \ar[r, "h_1"] &(\E :\B)^\Delta \ar[rd, phantom, very near start, "\lrcorner"] \ar[r, "h_2"] \ar[d, "\cod" description]
  & (\E:\B)^\to \ar[d, "\cod"] \\
\B\ar[r, "1" swap] &\B \ar[r, "k" swap] & \E:\B\\
\end{tikzcd}
\]
is a pullback, where $h$ takes $E$ to $[\rid_{E}, E, \rid_{pE}]$,
and $k$ takes $B$ to $B$.

Suppose $\C$ is a category and $F : \C \to (\E : \B)^\Delta$
and $G : \C \to \B$ are functors, such that
$\cod \o F = k \o G$. So $F(C)$ is a diagonal morphism;
the universal map takes $C$ to the domain of that morphism.
\cqed
\end{proof}

\section{Exercises}

\begin{lemma}
Let $F_0 : n \to M_0$ be given.
Mesh morphisms
$$(F_0, F_1) : \sharp n \to (M_0, M_1, \del_M)$$
correspond bijectively to objects in $M_1$
 over $F_0(0) \times \cdots \times F_0(n-1)$.
\end{lemma}

\begin{proof}
Given $F_1 : \sharp n \to M_1$, then
$$F_1 (\theta, (F_0(0) \times \cdots \times F_0(n-1)))$$
is the required element of the fiber.
Conversely, given an object $X$ of $M_1$ over
$(F_0(0) \times \cdots \times F_0(n-1))$.
Want to construct an $F_1 : \sharp n \to M_1$.
\[
\begin{tikzcd}
\sharp n \ar[rd, phantom, very near start, "\lrcorner"] \ar[r, ""] \ar[d, "\del" swap]  & \R_n^\to\ar[d, "\cod"]  & M_1 \ar[d, "\del_M"]\\
P(n) \ar[rr, bend right, "P(F_0)" swap] \ar[r, "\iota" swap, hook] & \R_n \ar[ur, dotted] & P(M_0)
\end{tikzcd}
\]
{\color{red} XXX WRONG The dotted arrow is produced by the universal property of $\R_n$?}


\cqed
\end{proof}

\section{Generally}
Suppose I have a category $\C$ with finite products and I have a pullback
\[
\begin{tikzcd}
 M(\C) \ar[rrr, bend left] \ar[rd, phantom, very near start, "\lrcorner"]
\ar[r, ""] \ar[d, "\del" swap]  & \C^\to\ar[d, "\cod"] & & M_1 \ar[d, "\del_M"]\\
P(n) \ar[rrr, bend right, "P(f)" swap] \ar[r, "\iota" swap, hook] & \C \ar[r] & M_1 : P(M_0) & P(M_0)
\end{tikzcd}
\]

I want to think that $M(\C)$ is somehow the best approximation of $\C$ as a c-mesh.

Under what circumstances can I lift $P(f)$? For each object
 $m_0 \times \cdots \times m_{n-1}$ in $P(n)$ and each
$C \to D_{m_0} \times \cdots \times D_{m_{n-1}}$ in $\C$,
I need to produce an object of $M_1$ over $f(m_0) \times \cdots f(m_{n-1})$.

Might as well let $n$ be 1. For each basic object $m$ in $P(n)$,
and each $C \to D_m$ in $\C$, I need to produce an object of $M_1$ over
$f(m)$.
\section{More Specifically}
Given an object $X$ of $M_1$ over $x\in M_0$, construct an
$F_1 : \sharp 1 \to M_1$ over $P(x) : P(1) \to P(M_0)$.
\[
\begin{tikzcd}
\sharp 1 \ar[rd, phantom, very near start, "\lrcorner"] \ar[r, ""] \ar[d, "\del" swap]  & \R_1^\to\ar[d, "\cod"]  & M_1 \ar[d, "\del_M"]\\
P(1) \ar[rr, bend right, "P(x)" swap] \ar[r, "\iota" swap, hook] & \R_1 \ar[ur, dotted] & P(M_0)
\end{tikzcd}
\]

$\R_1$ is the free category with finite products
 on two objects $R$ and $Z$ and a morphism $R \to Z$.

\[
\begin{tikzcd}
\sharp n
    \ar[rd, phantom, very near start, "\lrcorner"]
    \ar[r, ""]
    \ar[d, "\del" swap]
  & \R_n^\to\ar[d, "\cod"] \ar[r, dotted]
  & Q^\to \ar[from=r] \ar[d, "\cod"]
  & M_1 \ar[ld, phantom, very near start, "\llcorner"] \ar[d, "\del_M"]
\\
P(n)
    \ar[r, "\iota" swap, hook]
  & \R_n \ar[r, dotted]
  & Q \ar[from=r, "\iota"]
  & P(M_0)
\end{tikzcd}
\]


What's $Q$? Its objects are the objects of $M_1$ and $P(M_0)$.
We write $[m]$ for an object derived from an object $m \in M_1$.
An object $m \in M_1$ becomes an arrow of $Q$ from $[m] \to \del m$


\section{WTF}

Ok, what's going on? I want to find `the free thing in the category of
opfibrations' that has some property. I want to suspect that it looks
something like a codomain opfibration most of the time.

Like: what's the free opfibration with a chosen element $X$ upstairs
over a chosen element $Y$ downstairs?

\[
\begin{tikzcd}
 \E_1 \ar[rrr, bend left, dotted] \ar[rd, phantom, very near start, "\lrcorner"]
\ar[r, "e"] \ar[d, "p_1" swap]  & \C^\to\ar[d, "\cod"] & & \E_2 \ar[d, "p_2"]\\
\B_1 \ar[rrr, bend right, "b" swap] \ar[r, "h"]
    & \C \ar[r, "k"] & \E_2 : \B_2 & \B_2 \ar[l, hook]
\end{tikzcd}
\]
objects of $\E_1$ are pairs $(B_1, (f : C \to  h(B_1)))$. We have that
$k(h(B_1)) = b(B_1)$.

So $k$ sends an object of $\E_1$ to $b(B_1)$.

We can hope that $k \o \dom$ is always in $E_2$, and that's what we pick?

\section{WTF2}

How about something easier; Let's just prove the free mesh on a set
functions correctly.

Mesh morphisms out of the free $c$-mesh on a set $X$
correspond to a set function $X \to M_0$.
\[
\begin{tikzcd}
 P(X)^\to \ar[r, "F_1"]\ar[d, "\cod" swap] & M_1 \ar[d, "\del"]\\
 P(X) \ar[r, "F_0" swap] & M_0
\end{tikzcd}
\]
Suffices that they correspond to product-preserving functors $P(X) \to M_0$
because we just lean on the freeness of $P$ after that.

\subsection{Suffices to show}
Let $(M_0, M_1, \del)$ be a mesh. Mesh morphisms from
the codomain opfibration on a category $\C$ with finite products
to $(M_0, M_1, \del)$
correspond to finite-product-preserving functors $\C \to M_0$.
\[
\begin{tikzcd}
 \C^\to \ar[r, "F_1"]\ar[d, "\cod" swap] & M_1 \ar[d, "\del"]\\
 \C \ar[r, "F_0" swap] & M_0
\end{tikzcd}
\]
So let an $F_0$ be given, we then have to construct $F_1$. Given
$f : C \to D \in \C$ we want to find $f^\star \in M_1$ such that
$\del f^\star = F_0 D$. But $\del$ is a fibration, so we can pull $F_0 D$
back along $F_0f : F_0C \to F_0 D$ to get exactly the arrow in $M_1$
we need.
\end{document}
