\documentclass{article}
\input{theorem}
\input{prooftree}
\usepackage{stmaryrd}
\input{linear}
\def\//{\mathop{/\!\!/}}
\def\div{\vskip 1em \hrule \vskip 1em}
\def\pbck{\ar[dr, phantom, pos=0, "\lrcorner"]}
\def\bpbck{\ar[ddrr, phantom, pos=0, "\lrcorner"]}
\def\lpbck{\ar[dl, phantom, pos=0, "\llcorner"]}
\def\erule#1#2{\begin{prooftree}#1\justifies #2\end{prooftree}}
\def\lpar{\bindnasrepma}
\def\lamp{\binampersand}
\def\bpush{\mathbf{push}}
\newarrow {Equiv} =====
\def\FinSet{\mathbf{FinSet}}
\def\x{\times}
\def\o{\mathrel\comp}
\def\ups#1{#1^G}
\def\dns#1{#1^B}
\def\rsn{\ \mathsf{sn}}
\def\rx{\mapsto}
\def\ep{\varepsilon}
\def\wat{{@}}
\def\sem#1{[\![#1]\!]}
\def\cn{{:}}
\def\rok{\mathrel\mathsf{ok}}
\def\rtype{\mathrm{type}}
\def\rset{\mathbf{Set}}
\def\rfinset{\mathbf{FinSet}}
\def\rspan{\mathbf{Sp}}
\def\rmesh{\mathbf{Mesh}}
\def\rhgrph{\mathbf{HGrph}}
\def\b2{\mathbf 2}
\def\rcat{\mathbf{Cat}}
\def\rcart{\mathbf{Cart}}
\def\rfincat{\mathbf{FinCat}}
\def\rid{\mathrm{id}}
\def\maps{\ar[dd,mapsto,shorten <=2em, shorten >=2em]}
\def\mapt{\ar[ddd,mapsto,shorten <=2em, shorten >=2em]}
\def\frommaps{\ar[from=uu,mapsto,shorten <=2em, shorten >=2em, crossing over]}
\def\frommapt{\ar[from=uuu,mapsto,shorten <=2em, shorten >=2em, crossing over]}
\def\C{\mathbf{C}}
\def\D{\mathbf{D}}
\def\J{\mathbf{J}}
\def\R{\mathbf{R}}
\def\E{\mathbf{E}}
\def\F{\mathbf{F}}
\def\T{\mathbf{T}}
\def\B{\mathbf{B}}
\def\M{\mathbf{M}}
\def\H{\mathbf{H}}
\def\pair#1#2{\langle#1,#2\rangle}
\def\tri{\triangleright}
\def\o{\circ}
\def\rctx{\,\mathrm{ctx}}
\def\del{\partial}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\def\also#1{\ \textcolor{blue}{\celse #1}}
\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{positioning}
\newcounter{nodemaker}
\setcounter{nodemaker}{0}
\def\twocell#1#2{%
  \global\edef\mynodeone{twocell\arabic{nodemaker}}%
  \stepcounter{nodemaker}%
  \global\edef\mynodetwo{twocell\arabic{nodemaker}}%
  \stepcounter{nodemaker}%
  \ar[#1,phantom,shift left=3,""{name=\mynodeone}]%
  \ar[#1,phantom,shift right=3,""'{name=\mynodetwo}]%
  \ar[Rightarrow,#2,from=\mynodeone,to=\mynodetwo]%
}
\def\twocellswap#1#2{%
  \global\edef\mynodeone{twocell\arabic{nodemaker}}%
  \stepcounter{nodemaker}%
  \global\edef\mynodetwo{twocell\arabic{nodemaker}}%
  \stepcounter{nodemaker}%
  \ar[#1,phantom,shift right=3,""{name=\mynodeone}]%
  \ar[#1,phantom,shift left=3,""'{name=\mynodetwo}]%
  \ar[Rightarrow,#2,from=\mynodeone,to=\mynodetwo]%
}

\def\maps{\ar[dd,mapsto,shorten <=2em, shorten >=2em]}
\author{Jason Reed}
\definecolor{fib}{HTML}{ff7f00}
\definecolor{opfib}{HTML}{007fff}
\definecolor{dgreen}{HTML}{005f00}
\definecolor{mild}{HTML}{cc4444}

\begin{document}

\tikzset{
   commutative diagrams/.cd,
   arrow style=tikz,
   diagrams={>=stealth}}
% got this from
% https://tex.stackexchange.com/questions/169512/tikz-style-arrow-tips-missing-when-using-tikz-cd-crossing-over
% ???


\section{Meshes}
\subsection{Preliminary Definitions}
Where by $L(\C)$ I mean `the free finitely complete category on $\C$'.
When $f : \C \to \D$ and $\D$ a finitely complete category,
write $\bar f$ for the unique map $L(\C) \to \D$. For an object $\ell \in L(C)$,
we write $\dom \ell$ for the `index set of $\ell$', thinking of $\ell$ as an object of
 $(\rfinset / \C)^\op$.

\subsection{Defining Meshes}
There is a category $\rmesh$ whose objects are functions
$X \to \D$ where $\D$ is finitely complete, and whose morphisms
are squares
\[
\begin{tikzcd}
\D_1\ar[r, "d"]\ar[from=d]& \D_2\ar[from=d]\\
X_1\ar[r, "x" swap]& X_2
\end{tikzcd}
\]
such that $d$ preserves all finite limits.
Some examples of meshes are
\[
\begin{tabular}{lll}
Intuition&$X$&$\D$\\
\hline
Walking $n$-ary Relation&$n $&$ L(\pi_i : R \to n)$\\
Set $X$&$X $&$ L(X)$\\
Category $\C$&$\C_0 $&$ L(\C)$\\
\end{tabular}
\]


The `objects' of a mesh $M : X \to \D$ are elements of $X$. The
things that are allowed to be `the boundary of a morphism' are objects
of the category $L(X)$, i.e. lists of elements of $X$, construed as a product.
The notion of `boundary' is analogous to the pair of
the domain and codomain of a morphism in a category
--- except of course that the boundary might not be two objects, but instead
fewer or more. For
some $x \in L(X)$, the things we interpret as `morphisms on the boundary $x$' are
morphisms $D \to M x \in \D$. We take products of relations by
just taking products of morphisms in $\D$. Just like in the codomain
bifibration, we do fibrational pulls in $\D$ by pullback and opfibrational
pushes by postcomposition.

\subsection{Hypergraphs}

Categories are to graphs as meshes are to hypergraphs --- a hypergraph is what
you get if you strip any data about how `composition' works from a mesh.
We'll want to talk about the free mesh on a hypergraph the same way we'd talk
about the free category on a graph.

A hypergraph is a span of sets
\[
\begin{tikzcd}
& P\ar[dl, "\rho" swap]\ar[dr, "\theta"]\\
R&& Q
\end{tikzcd}
\]
where $R$ should be thought of as the relations of the premesh,
$Q$ as objects of the premesh, and the set $P$ consisting of all the projections
out of all relations to their objects. The category of these is just
$\rhgrph = \rset^{\rspan}$.

I want to now describe the free-forgetful adjunction
\[
\begin{tikzcd}
 \rhgrph \ar[r, shift left, "F"]\ar[from=r, shift left, "U"]&\rmesh
\end{tikzcd}
\]
The underlying hypergraph $UM$ of a mesh $M : X \to \C$ is defined like so:
\begin{itemize}
\item $Q$ is $X$
\item $R$ is the set of tuples $(C \in \C_0, \ell \in (L(X))_0,  \rho : C \to \bar M \ell)$, and
\item $P$ is the set of tuples
$$(C \in \C_0, \ell \in (L(X))_0,  \rho : C \to \bar M \ell, \delta \in \dom \ell)$$
The projection from $P$ to $R$ is the evident projection away from $\delta$,
and the function from $P$ to $Q$ takes $(C, \ell, \rho, \delta)$ to $M (\ell \delta)$.
\end{itemize}

The free mesh $F \Gamma$ of a hypergraph $\Gamma$ is defined like so:
\begin{itemize}
\item $X$ is $Q$
\item $\D = L(\Gamma)$ where we interpret $\Gamma$ as a category
whose objects are $R \cup Q$, and which has one morphism for each $p \in P$,
going from $\rho(p)$ to $\theta(p)$.
%% \item The objects of $\D$ are ways of filling in the diagram
%% \[
%% \begin{tikzcd}
%% X\ar[d]\ar[from=r]& Y\lpbck\ar[r]\ar[d]& Z\ar[d]\ar[from=r]& W \ar[dl, bend left]\\
%% R\ar[from=r]& P \ar[r] &Q
%% \end{tikzcd}
%% \]
\end{itemize}

We should confirm that $F \Gamma \to M$ actually does correspond to $\Gamma \to UM$
as expected.

On both sides, objects go to objects.
On the left, we have to map the category $\Gamma$ into $M$, preserving
what we do to objects in $Q/X$. So we have a choice of where
to send all the things in $R$, but we have to pick maps for all the projections ---
and the projections have to go into {\em objects}. But on the right, we picked
a description of $R$ that matches this. So I think this is right.

\subsection{Checking Conditions}
We need $UF : \rhgrph \to \rhgrph$ to be a coproduct
of representables. Let $[0]$ be
\[
\begin{tikzcd}
& 0\ar[dl, "!" swap]\ar[dr, "!"]\\
0&& 1
\end{tikzcd}
\]
And we can see that $(UF(X))_Q = \hom([0], X)$. Now
we need to come up with a bunch of hypergraphs that characterize the situations
where relations are required to exist. Any single $n$-ary relation
\[
\begin{tikzcd}
& n\ar[dl, "!" swap]\ar[dr, equal]\\
\{r\}&& n
\end{tikzcd}
\]
might have come from the product of two smaller relations. So any
hom from
\[
\begin{tikzcd}
& P\ar[dl, "\rho" swap]\ar[dr, "\theta"]\\
\{0, 1\}&& Q
\end{tikzcd}
\]
is going to imply another cell exists.

Actually, wait... doesn't any hom from anything at all of the form
\[
\begin{tikzcd}
& P\ar[dl, "\rho" swap]\ar[dr, "\theta"]\\
R&& Q
\end{tikzcd}
\]
determine a generated relation?

Yeah, they actually determine a bunch: such a span tells me which relations
I want to product together, and how I want to bind their object bits together
(or indeed which remain unbound? yeah! objects of $Q$ not in the image of
anything) but it doesn't tell me how to push anything --- but that's ok,
because I know all the pushes I might do, irrespective of how the hom comes in.

So I think the answer to `what is the corresponding thing to $\Delta$ in my case'
is {\em all} of the finitely generated free meshes.

\subsection{Reducing this to something Manageable}

I'd like to hope that the category --- let's call it $\Phi$ or
something --- is itself describable in some nice finite way.
%% What are
%% the four operations that I think of as fundamental to building up the free
%% mesh on a hypergraph?

%% \[
%% \begin{tabular}{ll}
%% Pull&$X\ X \to X$\\
%% Push&$ X \to X\ X$\\
%% Pull&$1 \to X$\\
%% Push&$ X \to 1$\\
%% \end{tabular}
%% \]

Let $T = UF$ be the {\em free mesh monad} on the category of hypergraphs.
We can directly describe what it means to be a function $\Gamma_1 \to T\Gamma_2$,
which is the same thing as a map between free meshes.

We reason through this in a series of steps. There are three basic sorts of maps
$\Gamma_1 \to T\Gamma_2$ we can build up all the rest from. The first is
a {\em product map}
\[
\begin{tikzcd}
R_1\ar[from=r]
  &P_1  \ar[d, equal]\ar[r]
  & Q_1 \ar[dd, equal]\\
R_0 \ar[u]\ar[d]
  & P_1 \lpbck \ar[l]\ar[d]\\
R_2 \ar[from=r ]& P_2\ar[r]& Q_2
\end{tikzcd}
\]
In which every relation $r \in R_1$ is mapped to a product of relations
in $R_2$. The second is a {\em pull map}
\[
\begin{tikzcd}
R_1\ar[from=r] \ar[d, equal]
  &P_1  \ar[from=d]\ar[r]
  & Q_1 \ar[d, equal]\\
R_2 \ar[from=r ]& P_2\ar[r]& Q_2
\end{tikzcd}
\]
in which relations in $R_1$ have some of their projections pulled together.
The third is a {\em push map}
\[
\begin{tikzcd}
R_1\ar[from=r] \ar[d, equal]
  &P_1  \ar[d]\ar[r]
  & Q_1 \ar[d]\\
R_2 \ar[from=r ]& P_2\ar[r]& Q_2
\end{tikzcd}
\]


Every map in the category of free meshes on a hypergraph is a product followed
by a pull followed by a push, so a general map looks like
a choice
of $X, Y, z_1, z_2, r_0, r_1, r_2, q_0, q, y_1, y_2$ such that the diagram
%% \[
%% \begin{tikzcd}
%% R_1 \ar[from=rr, "\rho_1" swap]
%%   &
%%   & P_1 \ar[d, "\pi_1"]\ar[r, "\theta_1"]
%%   & Q_1 \ar[dd, "q"]\\
%% X \ar[from=r]\ar[u, "r_1"]\ar[d, "r_2" swap]
%%   & X \x_{R_2} P_2 \ar[r, "\pi_2" swap]\ar[d, "\rho_2^*r_2"] \lpbck
%%   &Y \ar[dr, "y" description]\ar[ull, "\phi" description]\\
%% R_2 \ar[from=r, "\rho_2" ]& P_2\ar[rr, "\theta_2" swap]&& Q_2
%% \end{tikzcd}
%% \]
%% commutes.
\[
\begin{tikzcd}
R_1\ar[from=r, "\rho_1" swap]\ar[d, equal]
  &P_1  \ar[d, "y_1"]\ar[r, "\theta_1"]
  & Q_1 \ar[d, "q"]\\
R_1
  &Y\ar[l, "r_0" description]\ar[from=d, "y_2" swap]\ar[r, "q_0" description]
  &  Q_2 \ar[dd, equal]\\
X \ar[from=r]\ar[u, "r_1"]\ar[d, "r_2" swap]
  & X \x_{R_2} P_2 \ar[d] \lpbck
  \\
R_2 \ar[from=r, "\rho_2" ]& P_2\ar[r, "\theta_2" swap]& Q_2
\end{tikzcd}
\]
commutes.

Example:
\[
\begin{tikzcd}[column sep=0.5em]
\{R\} \ar[from=rr, "" swap]
  &
  & \{R_a,R_b\} \ar[d, ""]\ar[r, ""]
  & \{a,b\} \ar[dd, ""]\\
\{ S,  T\} \ar[from=r]\ar[u, ""]\ar[d, "" swap]
  & \{S_c, S_d, T_d, T_e\} \ar[r, ""]\ar[d, ""] \lpbck
  &\{c,d,e\} \ar[dr, ""]\\
\{S, T\} \ar[from=r, "" ]& \{S_c,S_d,T_d,T_e\}\ar[rr, "" swap]&& \{c,d,e\}
\end{tikzcd}
\]

\[
\begin{tikzcd}
R_1
  &Y\ar[l, "r_0" description]\ar[from=d, "y_2" swap]\ar[r, "\theta_0" description]
  &  Q_2 \ar[dd]\\
X \ar[from=r]\ar[u, "r_1"]\ar[d, "r_2" swap]
  & Y \ar[d]
  \\
R_2 \ar[from=r, "\rho_2" ]& P_2\ar[r, "\theta_2" swap]& Q_2
\end{tikzcd}
\]

\vfil\eject
\section{Not Sure Above Stuff is Right}

Let me revisit the free-forgetful adjunction, as I'm not sure I got it right.
A hypergraph has vertices, edges, and projections.
\[
\begin{tikzcd}
& P\ar[dl, "\ep" swap]\ar[dr, "\nu"]\\
E&& V
\end{tikzcd}
\]
And maps in $\rhgrph$ are just span morphisms.

A mesh is a map $M : X \to \D$ with $\D$ finitely complete.
Maps in $\rmesh$ are commutative squares with one side preserving finite
limits.

There are a whole bunch of adjoints
\[
\begin{tikzcd}[row sep=3em]
\rset \ar[r, "F_\rhgrph", shift left]\ar[from=r, "U_\rset", shift left]
&\rhgrph \ar[r, "F_\rcat", shift left]\ar[from=r, "U_\rhgrph", shift left]
&\rcat \ar[r, "L", shift left]\ar[from=r, "U_\rcat", shift left]
&\rcart\\
&\rmesh \ar[u, "U_\rhgrph", shift left, near start]
        \ar[from=u, "F_{\rmesh}", shift left, near end]
\end{tikzcd}
\]
We drop subscripts on functors when unambiguous.

The free hypergraph $F(X)$ of a set $X$ is
\[
\begin{tikzcd}
& 0\ar[dl, "!" swap]\ar[dr, "!"]\\
0&& X
\end{tikzcd}
\]
The underlying set $U(H)$ of a hypergraph is its set of vertices.

The free category $F(H)$ of a hypergraph has objects $E \cup V$ and
(nonidentity) morphisms $P$, with $\dom := \ep$ and $\cod := \nu$
with no nontrivial compositions being possible. The underlying hypergraph
$U(\C)$ of a category has $E = V = \C_0$ and $P = \C_1$ with
$\ep := \dom$ and $\nu := \cod$.

The free finitely complete category on a category $\C$ is written $L(\C)$.

The free mesh of a hypergraph $F(H)$ has $X := V$ and $\C = L(F_{\rcat}(H))$,
with the injection arising from the evident function $V \to F_{\rcat}(H)$.
The underlying hypergraph of a mesh $U(M)$ is
\[
\begin{tikzcd}
& \C / M \ar[dl, "\dom" swap]\ar[dr, "\cod"]\\
\C_0 && X
\end{tikzcd}
\]

\begin{lemma}
$F_\rmesh \dashv U_\rhgrph$
\end{lemma}

\begin{proof}
Let $H \in \rhgrph$ and $M \in \rmesh$ be given.
Hypergraph morphisms $H \to UM$ are the same thing as a span map
\[
\begin{tikzcd}
E \ar[from=r]\ar[d]& \ar[r]\ar[d]P & V\ar[d]\\
\C_0 \ar[from=r]& \ar[r]\C/M & X
\end{tikzcd}
\]
A mesh morphism $F_\rmesh H \to M$ can be seen to be the same thing;
we have to map $V \to X$ downstairs, and we have to map
$L(F_{\rcat}(H)) \to M$ upstairs. This is the same thing as mapping
$F_{\rcat}(H) \to M$ upstairs. The objects of $F_{\rcat}(H)$ are $V
\cup E$, but we've already been told how we must map $V$, so we get a
choice of function $E \to \C_0$. The morphisms of
$F_{\rcat}(H)$ have to get sent to things in $\C/M$ because all the vertices
are pinned to $X$, and domains and codomains must be consistent with existing maps.
 \cqed
\end{proof}

\subsection{Characterizing Maps between Free Meshes}

Abbreviate $\M = F_\rmesh$.
Let a hypergraph map $f : H_1 \to \M(H_2)$ be given. What can it look like?
The category of $H_2$ is $\C(H_2)$. The injection of $V_2$ into $L(\C(H_2))$
let's call $\iota$. It's gotta be a span map
\[
\begin{tikzcd}
E_1 \ar[from=r]\ar[d]& \ar[r]\ar[d]P_1 & V_1\ar[d]\\
L(\C(H_2))_0 \ar[from=r, "\dom"]& \ar[r, "\cod" swap]L(\C(H_2))/\iota & V_2
\end{tikzcd}
\]

What's an object of $L(\C)$? A functor $\C \to \rset$. Or, equivalently,
a discrete fibration $\J \to \C^\op$. Or again equivalently in the special
case where $\C$ is $\C(H)$ for some $H = (E, P, V)$, merely a choice of sets
$(E', P', V')$ and
\[
\begin{tikzcd}
E' \ar[from=r]\ar[d]& \ar[r]\ar[d]P' & V'\ar[d]\\
E \ar[from=r]& \ar[r]P & V
\end{tikzcd}
\]

What's an object of $L(\C)$? A functor $\C \to \rset$. What's a morphism of $L(\C)$? A natural transformation going backwards.
So given $H_2$ looking like
\[
\begin{tikzcd}[column sep=0]
& e_1\ar[dl]\ar[dr]&& e_2\ar[dl]\ar[dr]\\
v_1&&v_2&&v_3
\end{tikzcd}
\]
a morphism in $L(\C(H_2))/\iota$ (let's say its codomain is $V_3$) is a natural transformation
\[
\begin{tikzcd}[column sep=-1em]
& \{\cdots\}\ar[dl]\ar[dr]&& \{\cdots\}\ar[dl]\ar[dr]\\
\{\cdots\}&&\{\cdots\}&&\{\cdots\}
\end{tikzcd}
\quad \Leftarrow\quad
\begin{tikzcd}[column sep=0]
& 0\ar[dl]\ar[dr]&& 0\ar[dl]\ar[dr]\\
0&&0&&1
\end{tikzcd}
\]
So an object of $L(\C(H_2))$ is a choice of a set $E_\bullet$ of {\em edge points} (elements
the coproduct of all the edge sets) and a choice of a set $V_\bullet$ of {\em vertex points}
(elements of the coproduct of all the vertex sets) and some maps
\[
\begin{tikzcd}
E_\bullet  \ar[d] & V_\bullet\ar[d] \\
E_2 & V_2
\end{tikzcd}
\]
We also need to know how each edge point gets transported under various projections.
So the dotted map in
\[
\begin{tikzcd}
E_\bullet \ar[d] \ar[from=r]& \cdot \ar[d] \lpbck \ar[r, dotted] & V_\bullet \ar[d]\\
E_2 \ar[from=r] & P_2 \ar[r] & V_2
\end{tikzcd}
\]
must exist.

All of these sets are fibered over $E_1$, so we have to have maps
\[
\begin{tikzcd}
E_1 \ar[from=r]\ar[from=drr]\ar[from=d] & P_1 \ar[r] & V_1\\
E_\bullet \ar[d] \ar[from=r]& \cdot \ar[d] \lpbck \ar[r] & V_\bullet \ar[d]\\
E_2 \ar[from=r] & P_2 \ar[r] & V_2
\end{tikzcd}
\]
And the projections $P_1$ have to pick out elements of $V_\bullet$, and the $V_1$
have to pick out $V_2$ consistently, so we're left with
\[
\begin{tikzcd}
E_1\ar[from=r]\ar[d, equal]
  &P_1  \ar[d]\ar[r]
  & V_1 \ar[d]\\
E_1
  &V_\bullet\ar[l]\ar[from=d]\ar[r]
  &  V_2 \ar[dd, equal]\\
E_\bullet \ar[from=r]\ar[u]\ar[d]
  & \cdot \ar[d] \lpbck
  \\
E_2 \ar[from=r]& P_2\ar[r]& V_2
\end{tikzcd}
\]
isomorphic to what I conjectured before.
\div
More vertically compact:
\[
\begin{tikzcd}
E_1\ar[from=rr]
  &
  &P_1  \ar[d]\ar[r]
  & V_1 \ar[dd]\\
E_\bullet \ar[from=r]\ar[u]\ar[d]
  & \cdot \ar[d] \lpbck\ar[r]
  & V_\bullet \ar[dr] \ar[ull]
  \\
E_2 \ar[from=r]
 & P_2\ar[rr]
 &
 & V_2
\end{tikzcd}
\]
\div
What are some examples? Let's look at a binary pull, a map
\[
\begin{tikzcd}
R \ar[d] \\ a
\end{tikzcd}
\quad\to\quad
\M
\left(
\begin{tikzcd}
R \ar[d, bend right] \ar[d, bend left] \\ a
\end{tikzcd}
\right)
\]
Some typical objects of $\M(\cdots)$ are
\[
R =
\begin{tikzcd}
1 \ar[d, bend right, "\ell" swap] \ar[d, bend left, "r"] \\ \{\ell, r\}
\end{tikzcd}
\qquad
R \x R =
\begin{tikzcd}
2 \ar[d, bend right, "\ell" swap] \ar[d, bend left, "r"] \\ \{\ell_0, r_0, \ell_1, r_1\}
\end{tikzcd}
\]
\[
R \x a =
\begin{tikzcd}
1 \ar[d, bend right, "\ell" swap] \ar[d, bend left, "r"] \\ \{\ell, r, a\}
\end{tikzcd}
\qquad
R^\Delta =
\begin{tikzcd}
1 \ar[d, bend right] \ar[d, bend left] \\ 1
\end{tikzcd}
\]

\div
What's a better way of thinking about a functor $\C \to \rset$ for the specific
case of $\C(H)$?

I should be able to Grothendieck it around and think of it as a choice
of amalgamated sets of all $E$-elements and all $V$-elements.

\section{Remedial Category Theory}

\begin{lemma}
Any limit is a product followed by an equalizer.
\end{lemma}

\begin{proof}
A limit diagram in $\C$ with shape $\D$
is a functor $\J : \C^\D$. For each $C\in \C$ there is a constant
functor $\Delta(C) : \C^\D$. This is actually the diagonal functor
$\Delta : \C \to \C^\D$. Limits are right adjoint to the diagonal
$\Delta \dashv \lim_{\leftarrow}$
$$\begin{prooftree}
\Delta^\D_\C(C) \to \J :  \C^\D
\Justifies
C \to  \lim_{\leftarrow}{}^\D_\C\, \J : \C
\end{prooftree}$$

A product is a limit with discrete shape $D$.
$$\begin{prooftree}
\Delta^D_\C(C) \to \J :  \C^D
\Justifies
C \to  \lim_{\leftarrow}{}^D_\C\, \J : \C
\end{prooftree}$$


\cqed
\end{proof}

\end{document}
