\documentclass{article}
\usepackage[tmargin=0.05in, bmargin=0.05in]{geometry}
\input{theorem}
\input{prooftree}
\usepackage{relsize}
\usepackage{stmaryrd}
\usepackage{latexsym}
\usepackage{yfonts}
\usepackage{amsmath}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{tikz}
\usetikzlibrary{calc,arrows,cd,decorations.pathreplacing}
\usetikzlibrary{decorations.pathmorphing}
\usepackage{tcolorbox}
\tcbuselibrary{breakable}
\usepackage{listings}
\lstset{basicstyle=\ttfamily\footnotesize,breaklines=true}

\def\lint{{\textstyle\int}}
\def\isopen{\textcolor{orange}{\mathlarger{\mathlarger{\mathlarger{\textbf ?}}}}}
\def\isok{\textcolor{green}{\mathlarger{\mathlarger{\mathlarger{\checkmark}}}}}
\def\isbad{\textcolor{red}{\mathlarger{\mathlarger{\mathlarger{\x}}}}}
\def\chave{\eqno \textcolor{blue}{\hbox{have}}}
\def\cset{\eqno \textcolor{green}{\hbox{set}}}
\def\cwant{\eqno \textcolor{orange}{\hbox{want}}}
\def\gol{\sqsubset}
\def\gor{\sqsupset}

\def\sp{\bullet}
\def\ot{\leftarrow}
\def\prequiv{\dashv\vdash}
\def\fdom{{\mathbf{d}f}}
\def\fcod{{\mathbf{e}f}}
\def\fidom{\mathbf{\bar d}f}
\def\ficod{\mathbf{\bar e}f}

\def\thn{\mathrel|}
\def\coe{\mathsf{coe}}
\def\bpush{\mathbf{push}}
\def\bpull{\mathbf{pull}}

\def\wo{\setminus}
\def\fA{\mathsf{A}}
\def\fC{\mathsf{C}}
\def\fX{\mathsf{X}}
\def\ff{\mathsf{f}}
\def\fm{\mathsf{m}}
\def\rspan{\mathsf{Span}}
% \def\El#1{\mathsf{El}({#1})}
\def\El#1{\ulcorner{#1}\urcorner}
\def\U{\mathsf{U}}
\def\uu{\mathsf{u}}
\def\uni{\U}
\def\mor#1#2{#1 \to \underline{#2}}

\def\ridp{\mathsf{idp}}
\def\ssem#1{\langle\!\langle{#1}\rangle\!\rangle}

\definecolor{lred}{rgb}{0.95,0.8,0.8}
\definecolor{cyan}{rgb}{0.80,0.95,0.95}
\definecolor{lorange}{rgb}{0.95,0.90,0.80}
\definecolor{bluegray}{rgb}{0.6,0.6,0.85}
\definecolor{lgray}{gray}{0.9}
\newtcolorbox{declbox}{colframe=lred,colback=lred,grow to right by=-10mm,grow to left by=-10mm,
boxrule=0pt,boxsep=0pt,breakable}
\newtcolorbox{thmbox}{colframe=cyan,colback=cyan,grow to right by=-10mm,grow to left by=-10mm,
boxrule=0pt,boxsep=0pt,breakable}
\newtcolorbox{defnbox}{colframe=lorange,colback=lorange,grow to right by=-10mm,grow to left by=-10mm,
boxrule=0pt,boxsep=0pt,breakable}
\newtcolorbox{tangentbox}{colframe=lgray,colback=lgray,grow to right by=-10mm,grow to left by=-10mm,
boxrule=0pt,boxsep=0pt,breakable}

\input{linear}
\def\rpath#1{[#1]}
\def\rfst{\mathsf{fst}\,}
\def\rsnd{\mathsf{snd}\,}
\def\bunfold{\mathbf{unfold}}
\def\bfold{\mathbf{fold}}
\def\susp#1{\langle {#1} \rangle}
\def\ssusp#1{\langle\!\langle {#1} \rangle\!\rangle}
\def\dto{\mathrel{\dot\to}}
\def\dns{{\downarrow}}
\def\ups{{\uparrow}}
\def\sprov{\Vdash}
\def\signat{{\color{red}\Sigma}}
\def\ep#1{\underline{#1}}
\def\wat{\mathbin{@}}
\def\wwat{\mathbin{\underline\wat}}
\def\cn{{:}}
\def\dv{{\div}}
\def\ww{{\mathsf w}}
\def\base{{\textfrak b}}

\def\munit{I}
\def\tt{{\mathsf t}}
\def\ii{{\mathbb I}}
\def\T{\mathbb{T}}
\def\tw{\mathsf{Tw}}
\def\sw#1{^{#1}_{#1}}
\def\TP{\mathbf{TPre}}
\def\pp{\textsf{\textbf p}}
\def\nn{{\mathsf n}}
\def\PP{\textsf{\textbf P}}
\def\NN{\textsf{\textbf N}}
\def\zz{{\mathsf z}}
\def\sem#1{[\![#1]\!]}
\def\usem#1{\lceil\!\!\lceil{#1}\rceil\!\!\rceil}
\def\zem#1{\langle\!\langle#1\rangle\!\rangle}
\def\col#1{{\mathsf{C}}_{#1}}
\def\lax{\mathop\bigcirc}
\def\rjust{\mathsf{just}}
\def\rcvt{\mathsf{cvt}}
\def\rtype{\mathsf{type}}
\def\rdtype{\mathsf{disc}}
\def\rkind{\mathsf{kind}}
\def\rmode{\mathsf{mode}}
\def\rprel{\mathbf{Prel}}
\def\rrfam{\mathbf{Rfam}}
\def\rset{\mathbf{Set}}
\def\rfinset{\mathbf{FinSet}}
\def\rrel{\mathbf{Rel}}
\def\rcat{\mathbf{Cat}}
\def\rfam{\mathbf{Fam}}
\def\binj{\mathbf{inj}}
\def\rid{\mathsf{id}}
\def\rtrue{\mathrel\mathrm{true}}
\def\pbck{\ar[dr, phantom, pos=0, "\lrcorner"]}
\def\bpbck{\ar[ddrr, phantom, pos=0, "\lrcorner"]}
\def\erule#1#2{\begin{prooftree}#1\justifies #2\end{prooftree}}
\def\lpar{\bindnasrepma}
\def\lamp{\binampersand}
\def\btwo{\mathbf{2}}
\def\bone{\mathbf{1}}

\newarrow {Equiv} =====
\def\FinSet{\mathbf{FinSet}}
\def\x{\times}
\def\maps{\ar[dd,mapsto,shorten <=2em, shorten >=2em]}
\def\mapt{\ar[ddd,mapsto,shorten <=2em, shorten >=2em]}
\def\A{\mathbf{A}}
\def\B{\mathbf{B}}
\def\C{\mathbf{C}}
\def\G{\mathbf{G}}
\def\D{\mathbf{D}}
\def\DD{\mathbb{D}}
\def\E{\mathbf{E}}
\def\M{\mathbf{M}}
\def\pair#1#2{\langle#1,#2\rangle}
\def\ltri{{\lhd}}
\def\rtri{{\rhd}}
\def\tri{\rhd}
\def\ll{{<}}
\def\rr{{>}}
\def\cc{\mathsf{c}}
\def\dd{\mathsf{d}}
\def\o{\circ}
\def\rctx{\,\mathsf{ctx}}
\def\rdctx{\,\mathsf{dctx}}
\def\del{\partial}
\def\also#1{\ \textcolor{blue}{\celse #1}}
\newcounter{nodemaker}
\setcounter{nodemaker}{0}
\def\twocell#1#2{%
  \global\edef\mynodeone{twocell\arabic{nodemaker}}%
  \stepcounter{nodemaker}%
  \global\edef\mynodetwo{twocell\arabic{nodemaker}}%
  \stepcounter{nodemaker}%
  \ar[#1,phantom,shift left=3,""{name=\mynodeone}]%
  \ar[#1,phantom,shift right=3,""'{name=\mynodetwo}]%
  \ar[Rightarrow,#2,from=\mynodeone,to=\mynodetwo]%
}
\def\twocellswap#1#2{%
  \global\edef\mynodeone{twocell\arabic{nodemaker}}%
  \stepcounter{nodemaker}%
  \global\edef\mynodetwo{twocell\arabic{nodemaker}}%
  \stepcounter{nodemaker}%
  \ar[#1,phantom,shift right=3,""{name=\mynodeone}]%
  \ar[#1,phantom,shift left=3,""'{name=\mynodetwo}]%
  \ar[Rightarrow,#2,from=\mynodeone,to=\mynodetwo]%
}
\def\maps{\ar[dd,mapsto,shorten <=2em, shorten >=2em]}
\author{Jason Reed}
\definecolor{fib}{HTML}{ff7f00}
\definecolor{green}{HTML}{007f00}
\definecolor{red}{HTML}{df3f3f}
\definecolor{opfib}{HTML}{007fff}

% For lightening colors I used
% http://trycolors.com/ 6 parts white to 1 part color
\definecolor{opfibl}{HTML}{DBEDFF}
\definecolor{greenl}{HTML}{DBEDDB}
\definecolor{redl}{HTML}{FAE4E4}
\def\njudge#1{\fcolorbox{opfib}{white}{#1}}
\def\nrule#1{\fcolorbox{white}{opfibl}{#1}}
\def\neqn#1{\fcolorbox{white}{redl}{#1}}



\def\judge#1{\vskip 2em\noindent \njudge{$#1$}\vskip 0.5em\noindent}
\def\rule#1{\vskip 2em \noindent \nrule{$#1$}\vskip 0.5em\noindent}
\def\eqn#1{\vskip 2em \noindent \neqn{$#1$}\vskip 0.5em\noindent}
\def\prof{\nrightarrow}
\def\tensor{\otimes}
\def\htensor{\mathrel{\hat\otimes}}
\def\lol{\multimap}
\def\hlol{\mathrel{\hat\multimap}}
\def\wtf{{\color{red}???}}

\def\re{\mathbf{E}}
\def\sh{\sharp}
\def\shp{\mathrel{\sharp}}
\def\zero{\texttt{0}}
\begin{document}

\def\
\tikzset{
   commutative diagrams/.cd,
   arrow style=tikz,
   diagrams={>=stealth}}

\section{A Digression on Ordered Logic}

While I'm at FLoC, there's something not-very-main-work-related I'm
curious about. Suppose you take the language of ordered logic,
and, say that propositions are interpreted at
pairs of worlds, and

\[   (A \rrto B)(u, v) = \forall x . A(v, x) \to B(u, x)\]
\[   (A \llto B)(u, v) = \forall x . A(x, u) \to B(x, v)\]
\[   (A \bullet B)(u, v) = \exists x . A(u, x) \land B(x, v)\]

and contexts in the evident way interpreting comma as bullet,
as sequents \[\Omega \prov A\] as

\[ \forall uv . \Omega(u, v) \to A(u, v) \]

Do you then get exactly the proof theory of ordered logic? The
main suspicious thing I think I'd need to establish for completeness
is that when you end up with a state in which you have
$\Delta(u, v)$ on the left but also leftover junk from
previous applications of synchronous rules,
the leftover junk never helps you prove anything you shouldn't.

Soundness of this interpretation, on the other hand, should
be pretty easy.
\begin{lemma}
If \(\Omega \prov A\), then
\( \forall uv . \Omega(u, v) \to A(u, v) \).
\end{lemma}

\begin{proof}
Asynchronous rules pretty much come for free.
Let's look at the synchronous rules:

We can transform
\[
\erule
{\Omega_2 \prov A \qquad \Omega_1, B, \Omega_3 \prov C}
{\Omega_1, A \rrto B, \Omega_2, \Omega_3 \prov C}
\]
into
$\Omega(u, v), A(v, w) \prov B(u, w)$
\[
\erule
{\[
\[
\justifies
\Omega_2(w, y) \prov A(w, y)
\]
\[
\justifies
 \Omega_1(u, v),B(v, y), \Omega_3(y, z) \prov C(u, z)
\]
\justifies
\Omega_1(u, v),  A(w, y) \to B(v, y), \Omega_2(w, y), \Omega_3(y, z) \prov C(u, z)
\]}
{\Omega_1(u, v), \forall x . A(w, x) \to B(v, x), \Omega_2(w, y), \Omega_3(y, z) \prov C(u, z)}
\]
And $\llto$ and $\bullet$ go pretty similarly.
\cqed
\end{proof}

Aha! I see why completeness fails. This is untethered/simpson-style ordered logic, as opposed to tethered/pfenning-davies.
With additives, I can give a shibboleth proposition that distinguishes them:
\[A \llto B \rrto 0, B \prov (A \llto 0) \bullet B\]
Here's why it fails in real ordered logic. Assume $A$ and $B$ are positive atoms.
So focus on $A \llto B \rrto 0$ fails because we don't have $A$ yet. We must focus on the right, so we have
to consume the $B$. Then we have the goal $A, A \llto B \rrto 0 \prov 0$, and although we can capture the $A$,
we have no $B$.

However, we can prove

\[
\erule
{\[
\[
\[
\[
\[
\[
\[
\justifies
 B^v_w \prov B^v_w
\]
\[
\using \textcolor{blue}{0L}
\justifies
0^t_w \prov  0^t_v
\]
\justifies
 B^v_w \to 0^t_w, B^v_w \prov  0^t_v
\]
\using \textcolor{blue}{\forall L} \justifies
\forall x . B^v_x \to 0^t_x, B^v_w \prov  0^t_v
\]
\justifies
( B \rrto 0)^t_v, B^v_w \prov  0^t_v
\]
\[
\justifies
A^t_u \prov A^t_u
\]
\justifies
A^t_u, (A \llto B \rrto 0)^u_v, B^v_w \prov  0^t_v
\]
\justifies
(A \llto B \rrto 0)^u_v, B^v_w \prov (A \llto 0)^u_v
\]
\[
 \justifies B^v_w \prov B^v_w
\]
\justifies
(A \llto B \rrto 0)^u_v, B^v_w \prov (A \llto 0)^u_v \x B^v_w
\]}
{(A \llto B \rrto 0)^u_v, B^v_w \prov ((A \llto 0) \bullet B)^u_w}
\]

Where the critical inferences are the highlighted $0L$ (which
relies on the fact that
$0$ at any pair of worlds is the same) and the
$\forall L$, which is a  tethering-forbidden
choice of $x = w$ which is outside the interval $[t,v]$.

\subsection{Tethering}
\def\tok{\hat \#}
So how would I imagine the tethered version looks?
I can throw a linear token in the context that indicates
the current active range.
An unfocused sequent like
\[\Delta \prov A\]
will get interpreted as
\[\Delta^u_v,  \tok^u_v \prov A^u_v \]
So I can define shifts as
\[ (\ups P)^u_v = \exists t w . [u,v] \subseteq [t,w] \land  \tok^u_v \tensor P^u_v \]
\[ (\dns N)^u_v = \forall t w. [u,v] \subseteq [t,w] \to  \tok^t_w \lol P^{tu}_{wv} \]
and connectives like
\[ ( P \rrto N)^{tu}_{wv} = \forall x \in [v, w] . P^{tv}_{wx} \to  N^{tu}_{wx} \]
\[ ( P \llto N)^{tu}_{wv} = \forall x \in [t, u] . P^{tx}_{wu} \to  N^{tx}_{wv} \]
\[ (P \tensor P)^{tu}_{wv} = \exists x \in [u, v] . P^{tu}_{wx} \land P^{tx}_{wv} \]

Does this work? Let me do a toy derivation of $\ups (\dns N \tensor \dns N)$ or something.

\[
\erule
{\[
\justifies
\Delta^t_w \prov  (\dns N \tensor \dns N)^t_w
\]}
{\Delta^t_w, \tok^t_w \prov \ups (\dns N \tensor \dns N)^t_w }
\]
... wait I'm getting type errors all over the place.

Is it really only the conclusion truth judgment
that needs the four-argument translation?

Let's start over. Maybe positives take two and negatives take four.
We expect the translation of a stable sequent to look canonically like
\[[u, v] \subset [t, w], (\dns \Delta)^u_v,  \tok^t_w \prov (\ups P)^{tu}_{wv} \]
and allow permitting junk to exist outside of $[u, v]$ should be acceptable.

I try to define shifts as
\[ (\ups P)^{tu}_{wv} = \ups (\tok^t_w \tensor P^u_v) \]
\[ (\dns N)^u_v = \dns \forall t w. [u,v] \subset [t,w] \to  \tok^t_w \lol N^{tu}_{wv} \]
and connectives like
\[ ( P \rrto N)^{tu}_{wv} = \forall x \in [v, w] . P^{v}_{x} \to  N^{tu}_{wx} \]
\[ ( P \llto N)^{tu}_{wv} = \forall x \in [t, u] . P^{x}_{u} \to  N^{tx}_{wv} \]
\[ (P \tensor P)^{u}_{v} = \exists x \in [u, v] . P^{u}_{x} \land P^{x}_{v} \]
If I interpret $\subseteq$ as involving a transitive, nonreflexive relation $<$
between worlds, then I think I get the right behavior.

For what happens when we try to focus on $P \rrto N$? Our choice of
$t,w$ is forced by whatever $\tok$ currently is. We are appropriately
constrained to choose a proposition to focus on that's `in the
currently available window', though. Okay, at this stage I'm actually
focused on $(P \rrto N)^{tu}_{wv} = \forall x\in[v, w]. P^v_x \to N^{tu}_{wx}$.
I have to choose an $x$ that doesn't exceed or equal
$w$, and which also doesn't go `backwards in time'. Choosing the last
world before $w$ corresponds to grabbing the whole context to the
right. We now have to prove (in right focus) $P^v_x$ which should
correspond to actually carrying out the proof in the appropriate
subset of the context, since by the time we blur into $N$, the token
we throw back will be tailored to that subset. On the left
side, we have $N^{tu}_{wx}$, which is the right thing. The only dangling
thread is the fact that by the time we leave focus, there's extra junk in
the context that represented proof goals that we {\em used} to have.
So we end up with a situation like
\[
\begin{tikzcd}
  u \ar[r, "P \rrto N"] \ar[rr, bend right, "N"'] & v \ar[r, "P"]& w
\end{tikzcd}
\]
where we still have smaller steps between worlds labelled by old
inhabitants of the context, and bigger jumps labelled by `newer'
inhabitants, as we proceed up the deduction tree.

Can I inductively accommodate this by thinking about the completeness theorem like so?
Let $\Omega$ be a (source language) ordered logic context, and let $\Delta$ be a
(target language) context of world variables and translated propositions and such.
We define a simulation relation $\Delta \sim^u_v \Omega$ if there exists
a way of `reading' $\Delta$ from $u$ to $v$ as denoting $\Omega$.

\subsection{Digression}
Maybe I can get away with organizing things like
I try to define shifts as
\[ (\ups P)^{tu}_{wv} = \ups ([u,v] \subset [t,w] \land \tok^t_w \tensor P^u_v) \]
\[ (\dns N)^u_v = \dns (\forall t w. [u,v] \subset [t,w] \to  \tok^t_w \lol N^{tu}_{wv}) \]
and connectives like
\[ ( P \rrto N)^{tu}_{wv} = \forall x . P^{v}_{x} \to  N^{tu}_{wx} \]
\[ ( P \llto N)^{tu}_{wv} = \forall x . P^{x}_{u} \to  N^{tx}_{wv} \]
\[ (P \tensor P)^{u}_{v} = \exists x . P^{u}_{x} \land P^{x}_{v} \]
and maybe prove that
\[P^u_v \prequiv P^u_v \land u < v\]
\[N^{tu}_{wv} \prequiv t < u \to N^{tu}_{wv}\]
\[N^{tu}_{wv} \prequiv v < w \to N^{tu}_{wv}\]
which really just amounts to
\[P^u_v \prov u < v\]
\[t < u \to N^{tu}_{wv} \prov N^{tu}_{wv}  \]
\[v < w \to N^{tu}_{wv} \prov N^{tu}_{wv}  \]

Well, we can prove
\[\exists x . P^u_x\land P^x_v  \prov u < v\]
by i.h. and transitivity. How about
\[t < u \to (\forall x . P^{v}_{x} \to  N^{tu}_{wx}) \prov \forall x . P^{v}_{x} \to  N^{tu}_{wx}  \]
? By induction hypotheses, we could also equivalently try to prove
\[t < u \to (\forall x . P^{v}_{x} \to  N^{tu}_{wx}) \prov \forall x . t < u \to P^{v}_{x} \to  N^{tu}_{wx}  \]
Oh, this is the easy half. One of the two interacting cases is
\[v < w \to (\forall x . P^{v}_{x} \to  N^{tu}_{wx}) \prov \forall x . P^{v}_{x} \to  N^{tu}_{wx}  \]
where we use i.h. to prove equivalently
\[v < w \to (\forall x . P^{v}_{x} \to  N^{tu}_{wx}) \prov \forall x . v \le x \to x \le w \to P^{v}_{x} \to  N^{tu}_{wx}  \]
and we use transitivity to achieve that.

\subsubsection{Thinking about Completeness Proof}

\begin{lemma}[Completeness]\ \\
1. Suppose for all readings $\Delta\sim^u_v \Omega$ that we have $\Delta \prov [P^u_v]$.
Then $\Omega \prov [P]$.\\
2. Suppose for all pairs of readings $\Delta_1\sim^t_u \Omega_1$ and
$\Delta_2 \sim^v_w \Omega_2$
we have $\Delta_1, \Delta_2 [N^{tu}_{wv}]\prov Q^t_w$. Then $\Omega_1  [N] \Omega_2 \prov Q$.
\end{lemma}

\begin{proof}
We can prove (2) by induction on $N$. Suppose we're dealing with $P \rrto N$.
We get to assume that for any $\Delta_1, \Delta_2$ with $\Delta_1 \sim ^t_u \Omega_1$ and
some $\Delta_{2} \sim^v_w \Omega_2$ that we have
\[\Delta_1, \Delta_{2}[ (P\rrto N)^{tu}_{wv}] \prov Q^t_w\]
i.e.
\[\Delta_1, \Delta_{2}[ \forall x \in [v, w] . P^{v}_{x} \to  N^{tu}_{wx} ] \prov Q^t_w\]

The rule we want to aim to use is
\[
\erule
{\Omega_{2a} \prov [P] \qquad \Omega_1, [N], \Omega_{2b} \prov Q}
{\Omega_1[P\rrto N]\Omega_{2a}\Omega_{2b} \prov Q}
\]
so, towards using the i.h., we want to prove ... no, this isn't right.
\cqed
\end{proof}

Is it
\begin{lemma}[Completeness]\ \\
1. If there exists a reading $\Delta\sim^u_v \Omega$ such that $\Delta \prov [P^u_v]$,
then $\Omega \prov [P]$.\\
2. If $\Delta_1\sim^t_u \Omega_1$ and
$\Delta_2 \sim^v_w \Omega_2$ and
 $\Delta_1, \Delta_2 [N^{tu}_{wv}]\prov Q^t_w$, then $\Omega_1  [N] \Omega_2 \prov Q$.
\end{lemma}
\begin{proof}
\end{proof}
No, this isn't right either.

Maybe we can do
\begin{lemma}[Completeness]\ %
\begin{enumerate}
\item If $\Delta \prov [P^u_v]$, then there exists an $\Omega$ with $\Delta \sim^u_v \Omega$
and $\Omega \prov [P]$.
\item If $\Delta [N^{tu}_{wv}]\prov Q^t_w$ then at least one of the following is true:
\begin{enumerate}
\item there exist
  $\Omega_1$ and $\Omega_2$ such that
  \[\Delta \sim^t_u \Omega_1 \qquad \Delta \sim^v_w \Omega_2 \qquad \Omega_1  [N] \Omega_2 \prov Q\]
\item there exists $\Omega$ such that
$\Delta \sim^t_w \Omega$ and $\Omega \prov Q$
\end{enumerate}
\item If $\Delta, \tok^t_w \prov Q^t_w$, then
there exists $\Omega$ such that
$\Delta \sim^t_w \Omega$ and $\Omega \prov Q$.
\end{enumerate}
\end{lemma}
\begin{proof}
\item[Case:] $\rrto$\\
 We have a derivation of
\[
\erule
{\[
\Delta \prov [ P^{v}_{x}] \qquad \Delta[ N^{tu}_{wx} ] \prov Q^t_w
\justifies
\Delta[ P^{v}_{x} \to  N^{tu}_{wx} ] \prov Q^t_w
\] \qquad x \in [v, w]}
{\Delta[ \forall x \in [v, w] . P^{v}_{x} \to  N^{tu}_{wx} ] \prov Q^t_w}
\]
Using the induction hypothesis on $P$ and $N$, in case $(a)$ for $N$,
we find $\Omega_1, \Omega_2, \Omega_3$ such that
\[\Delta \sim^v_x \Omega_2 \qquad \Delta \sim^t_u \Omega_1  \qquad \Delta \sim^x_w \Omega_3\]
and
\[\Omega_2 \prov [P] \qquad \Omega_1  [N] \Omega_3\prov Q\]
so it's straightforward to derive
\[
\erule
{\Omega_{2} \prov [P] \qquad \Omega_1, [N], \Omega_{3} \prov Q}
{\Omega_1[P\rrto N]\Omega_{2}\Omega_{3} \prov Q}
\]
and output the contexts $\Omega_1$ and $\Omega_2, \Omega_3$ noting that we can
establish that $\Delta\sim^v_w \Omega_2, \Omega_3$.

In case $(b)$ for $N$, we immediately return the same reading.
\item[Case:] $\ups P$\\
\[
\erule
{\[
\justifies
\Delta, \tok^t_w, P^u_v \prov Q^t_w
\]}
{\Delta[ \ups (\tok^t_w \tensor P^u_v) ] \prov Q^t_w}
\]
By i.h., there is some reading of $(\Delta, P^u_v) \sim^t_w Q^t_w$.
It either uses $P^u_v$ or it doesn't. If it does, it breaks up
into two smaller readings $\Delta \sim^t_u \Omega_1$
and
$\Delta \sim^v_w \Omega_2$, and so we have $\Omega_1, P, \Omega_2 \prov Q$.
From that we can derive $\Omega_1[\ups P]\Omega_2 \prov Q$ as required.
Otherwise, it doesn't use $P^u_v$, and so we just know we have
a reading $\Delta \sim^t_w \Omega$ and $\Omega \prov Q$, which
also suffices for the conclusion.
\cqed
\end{proof}

\end{document}
