\documentclass{article}
\usepackage{amssymb}
\input{theorem}

\input{prooftree}
\def\erule#1#2{\begin{prooftree}#1\justifies #2\end{prooftree}}
\def\pair#1#2{\langle #1 , #2 \rangle}

\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{calc}
\definecolor{morange}{rgb}{1,0.56,0}
\definecolor{lorange}{rgb}{1,0.95,0.8}
\definecolor{mgreen}{rgb}{0,0.56,0}
\definecolor{lgreen}{rgb}{0.95,1,0.8}
\definecolor{mblue2}{rgb}{0,0.2,1.0}
\definecolor{lblue}{rgb}{0.8,0.95,1}
\definecolor{mred}{rgb}{0.9,0.1,0.1}
\definecolor{mgreen}{rgb}{0.1,0.5,0.1}
\definecolor{mblue}{rgb}{0.3,0.3,0.9}
\def\bitf#1{#1 [smooth, tension=0.8] coordinates {(-1.6,2) (-1,1) (0,0)}}
\def\bitg#1{#1 [smooth, tension=0.8] coordinates {(1,2) (0.55,1) (0,0)}}
\def\bitgrev#1{#1 [smooth, tension=0.8] coordinates {(0,0) (0.55,1) (1,2)}}
\def\bitfg#1{#1 [smooth, tension=0.8] coordinates {(0,0) (0,-2) }}
\def\binj{\mathbf{inj}}
\def\blet{\mathrel\mathbf{let}}
\def\bin{\mathrel\mathbf{in}}
\def\bmatch{\mathrel\mathbf{match}}
\def\bwith{\mathrel\mathbf{with}}
\def\pbck{\ar[dr, phantom, pos=0, "\lrcorner"]}
\def\pdbck{\ar[ddr, phantom, pos=0, "\lrcorner"]}
\def\ups{{\uparrow}}
\def\dns{{\downarrow}}
\def\adjust{\big|}
\def\O{\mathcal{O}}
\def\rid{\mathsf{id}}
\def\ridp{\mathsf{idp}}
\def\rcoe{\mathsf{coe}}
\def\rtype{\mathsf{type}}
\def\int{\square}
\def\bd{\partial}
\def\prov{\vdash}
\def\prequiv{\dashv\vdash}
\def\imp{\Rightarrow}
\def\cqed{\hskip2mm{\vrule width .5em height .5em depth 0em}} % at the end of a |P.
\def\o{\circ}
\def\lx{\bigcirc}
\def\B{\mathbf{B}}
\def\C{\mathbf{C}}
\def\D{\mathbf{D}}
\def\E{\mathbf{E}}
\def\R{\mathbb{R}}
\def\two{\mathbf{2}}
\def\S{\mathbb{S}}
\def\M{\mathbb{M}}
\def\X{\mathbf{X}}
\def\Y{\mathcal{Y}}
\def\x{\times}
\def\st{\mathrel|}
\def\rset{\mathbf{Set}}
\def\rcat{\mathbf{Cat}}
\def\op{\mathsf{op}}
\def\P{\mathbb{P}}
\def\I{\mathbb{I}}
\def\U{\mathbb{U}}
\def\N{\mathbb{N}}
\def\Z{\mathbb{Z}}
\def\tw{\mathbf{2}}
\def\dash{\hbox{---}}
\def\dom{\mathop{\mathrm{dom}}}
\def\cod{\mathop{\mathrm{cod}}}
\def\celse{\mathrel{|}}
\def\cn{{:}}
\def\rok{\mathrel\mathsf{ok}}
\def\llam#1{\langle {#1} \rangle}
\def\hf{{\odot}}

\begin{document}
\tikzset{>=stealth}
\tikzset{
   commutative diagrams/.cd,
   arrow style=tikz,
   diagrams={>=stealth}}

What is the `territory of sets'? I'd guess that its paths are all functors from the poset $[0,1]$
to $\rset$. I can similarly turn any category $\C$ into a territory by considering all functors $[0,1]$
to $\C$. Applying this construction to $[0,1]$ {\em doesn't} give me the representable interval $[0,1]$,
because it doesn't take into account continuity. It has {\em more} paths in it than the representable,
including discontinuous ones.

I might also want to impose {\em some} notion of continuity even on
other territories derived from categories, because I'm confused about
the following phenomenon. Let $f : a \to b $ be the morphism in the
walking arrow category $\two$. A functor $[0,1] \to \two$ is
determined by its fiber over $a$, which must be a (and can be any)
downward closed interval, so either $[0,x)$ (which is empty if $x=0$)
%
or $[0,x]$ for $x \in [0,1]$. I think of the map that sends $(0,x)$ to
$a$ is one where $f$ `takes place just before time $t$', and the map
that sends $[0,x]$ is one where $f$ `takes place just after time $t$'.
Call these $f^{t-}$ and $f^{t+}$. It seems clear that $f^{t-}$ and
$f^{u-}$ are in some sense `the same map' even for $t \ne u$ by
homotopical reparameterization of the interval, and likewise $f^{t+}$
and $f^{u+}$. But are $f^{t-}$ and $f^{t+}$ the same?

It seems difficult to make them so --- there just {\em does not exist}
a continuous monotone map from $[0,1]$ to itself fixing 0 and 1 that
turns $f^{t-}$ into $f^{t+}$. Possibly it would be an interesting
choice to convert $\C$ into a territory by breaking the symmetry and
including only $f^{t-}$s or $f^{t+}$s.

One way of expressing that requirement is in terms of one-sided limit topology, like so:
The territory derived from $\C$ has for paths the functors $F : [0,1] \to \C$
such that the fiber over any object $C$ is the union of of half-open
 intervals $(x,y]$ or the single point $\{0\}$.
 Symmetrically, I could have said $[x,y)$ or the single point $\{1\}$.

Let's consider the territory derived from $\rset$. Depict time as flowing from left to right,
as functions merge points together. I seem to have a choice between saying that the unique-up-to-homotopy
 function $2 \to 1$
can look like
\[(a)\quad \begin{tikzpicture}[baseline=-1.5em,
   dot/.style={circle,fill=black,inner sep=0pt, minimum size=5pt},
   circ/.style={circle,inner sep=0pt,minimum size=4pt,draw,fill=white}]
\begin{scope}[scale=1]
\draw plot [smooth, tension=0.8] coordinates {(-0.8,0.5) (-0.5,0.25) (0,0)};
\draw plot [smooth, tension=0.8] coordinates {(-0.8,-1.5) (-0.5,-1.25) (0,-1)};
\draw plot [smooth, tension=0.8] coordinates {(0.8,0) (0.5,-0.25) (0,-0.5)};
\node[circ]  at (0,0) {};
\node[circ]  at (0,-1) {};
\node[dot]  at (0,-0.5) {};
\end{scope}
\end{tikzpicture}
\quad
\hbox{or}
\qquad
(b)
\quad
\begin{tikzpicture}[baseline=-1.5em,
   dot/.style={circle,fill=black,inner sep=0pt, minimum size=5pt},
   circ/.style={circle,inner sep=0pt,minimum size=4pt,draw,fill=white}]
\begin{scope}[scale=1]
\draw plot [smooth, tension=0.8] coordinates {(-0.8,0.5) (-0.5,0.25) (0,0)};
\draw plot [smooth, tension=0.8] coordinates {(-0.8,-1.5) (-0.5,-1.25) (0,-1)};
\draw plot [smooth, tension=0.8] coordinates {(0.8,0) (0.5,-0.25) (0,-0.5)};
\node[dot]  at (0,0) {};
\node[dot]  at (0,-1) {};
\node[circ]  at (0,-0.5) {};
\end{scope}
\end{tikzpicture}
\]
I think $(a)$ is the better option, because it has the same number of points as the geometrically
glued-together space
\[
\begin{tikzpicture}[baseline=-1.5em,
   dot/.style={circle,fill=black,inner sep=0pt, minimum size=5pt},
   circ/.style={circle,inner sep=0pt,minimum size=4pt,draw,fill=white}]
\begin{scope}[scale=1]
\draw plot [smooth, tension=0.8] coordinates {(-0.8,0) (-0.5,-.25) (0,-0.5)};
\draw plot [smooth, tension=0.8] coordinates {(-0.8,-1) (-0.5,-0.75) (0,-0.5)};
\draw plot [smooth, tension=0.8] coordinates {(0.8,0) (0.5,-0.25) (0,-0.5)};
\node[dot]  at (0,-0.5) {};
\end{scope}
\end{tikzpicture}
\]
which I imagine as giving the overall behavior of the evolving function correctly.
So yeah, that breaks the symmetry. I think the inverse image of any object should
be an open set in the topology where basic open sets are half-open intervals
closed at the {\em bottom}, $[a,b)$ (plus the singleton $\{1\}$, which is a residue of
any basic open set $[1,x)$ coming from the lower-limit topology on $\R$)

This definition as such still seems a bit `evil' for $\rset$, but I
feel good about using the intuition from it to break symmetry for what to do
with smaller categories.

\subsubsection*{Fibrational Perspective}

To be less `evil' I might try to return to my thought from before and say that a
path in $\rset$ is a topological space $X$
together with a continuous map $X \to [0,1]$, such that every filling problem
\[
\begin{tikzcd}
  1 \ar[r, "x"]\ar[d, "{[0/i]}"'] & X \ar[d]\\
\I \ar[ur, dotted]\ar[r, "f"'] & \I
\end{tikzcd}
\]
(for $f$ monotone and continuous) is uniquely solvable in the category of topological spaces.
I {\em don't} expect this to be quite the same thing as a functor $[0,1] \to \rset$, because
the latter would allow situations like $(b)$ above. Let me try to see where the proof of their
equivalence fails, and maybe that will tell me a better, less evil way to characterize which functors
$[0,1] \to \rset$ are well-behaved. For I think `inverse image of equivalence classes of sets
are certain sorts of open intervals' doesn't even {\em work} --- for it can't detect
$(b)$-like patterns of {\em isomorphisms}.

Ok, so let's start with a fibration $p : X \to [0,1]$. From it we want to construct a functor
$F : [0,1] \to \rset$. The object part of $F$ is taking the fibers of $p$; we say
\[ F(t) = p^{-1}(t) = \{x \in X \st p(x) = t\}\]
The morphism part of $F$ is fibrant transport. If we know $t \le u$, then
\[ F(t \le u) : F(t) \to F(u) = \lambda x . \rcoe^{t \mapsto u}(x) \]
We're relying on the fact that for $t \le u$ we can come up with an $f : \I \to \I$ such
that $f(0) = t$ and $f(1) = u$. I want to convince myself that the choice of $f$ doesn't matter.
Let me write
\[
\erule
{x : X(f(0))}
{
i : \I \prov \rcoe_f(x, i) : X(f(i))
}
\]
Suppose I have a $g : \I \to \I$ that preserves $0$ and $1$.
There should be a {\em unique} $g^*$ that fills below:
\[
\begin{tikzcd}
1 \ar[rr, "x"]\ar[dr, "0" description]\ar[d, "0"'] &&    X \ar[d]\\[3em]
\I\ar[urr, dotted, "g^*" description]\ar[r, "g"']&\I \ar[ur, dotted, "f^*" description]\ar[r, "f"'] & \I
\end{tikzcd}
\]
but by diagram chasing, in fact $f^* \o g$ satisfies the conditions $g^*$ is supposed to satisfy,
so $f^* \o g = g^*$. Since $g$ preserves $1$, then $f^*(1) = f^*(g(1)) = g^*(1)$.


A similar argument shows that $F$ preserves composition.

In the other direction, let's start with a functor $F : [0,1] \to \rset$. How do we do the `Grothendieck
construction', i.e. construct
the total space $X$ over $[0,1]$? The carrier is the dependent pair type
\[X = (t : [0,1]) \x F(t)\]
and I think a basic open set of the topology is defined like so. For an open subset $U \subseteq [0,1]$
we define a PER $(t_1,x_1) \sim_U (t_2,x_2)$ in $X$ to hold if there exists $(t,x) \in X$ such that
\begin{enumerate}
\item $t_i \le t$
\item $[t_i,t] \subseteq U$
\item $F(t_i \le t)(x_i) = x$
\end{enumerate}
for $i = 1,2$.

A basic open set of $X$ is an equivalence class under $\sim_U$ for some $U$.

\begin{lemma}
The relation $\sim_U$ is symmetric and transitive.
\end{lemma}

\begin{lemma}
If $U \subseteq V$, and  $x \sim_U y$, then $x \sim_V y$.
\end{lemma}

\begin{lemma}
If $c \sim_U d$ and $c \sim_V d$, then $c\sim_{U \cap V} d$.
\end{lemma}

\begin{proof}
Let's break this up into
\[(d_t, d_x) \sim_U (c_t, c_x)\]
\[(d_t, d_x) \sim_V (c_t, c_x)\]
and ask for witnesses $t^*_U$ and $t^*_V$ and $x^*_U$ and $x^*_V$ such that
\[d_t \le t^*_U \qquad c_t \le t^*_U \qquad d_t \le t^*_V \qquad c_t \le t^*_V \]
\[[d_t, t^*_U] \subseteq U \qquad [c_t, t^*_U] \subseteq U \qquad [d_t, t^*_V] \subseteq V \qquad [c_t, t^*_V]\subseteq V \]
\[F(d_t \le t^*_U)(d_x) = x^*_U \qquad F(c_t \le t^*_U)(c_x) = x^*_U\]
\[F(d_t \le t^*_V)(d_x) = x^*_V \qquad F(c_t \le t^*_V)(c_x) = x^*_V\]
The time we want to advance to is $t^* := \min(t_U^*, t_V^*)$. We then check that
\[d_t \le t^* \qquad c_t \le t^*\]
\[[d_t, t^*] \subseteq U \cap V \qquad [c_t, t^*] \subseteq U \cap V\]
\[F(d_t \le t^*)(d_x) =  F(c_t \le t^*)(c_x) \]
to show that
\[(d_t, d_x) \sim_{U \cap V} (c_t, c_x)\]
\cqed
\end{proof}

\begin{lemma}
The intersection of two basic open sets is open.
\end{lemma}

\begin{proof}
Let $U, V \subseteq [0,1]$ be given, as well as $a_t \in U$ and $b_t \in V$
and $a_x \in F(a_t)$ and $b_x \in F(b_t)$. So we have two elements $a = (a_t,a_x)$ and $b = (b_t,b_x)$ from $X$.
These are the representatives of equivalence classes. We want to show that
\[[a]_{\sim_U} \cap [b]_{\sim_V}\]
is an open set. If it's empty, it is trivially open. If it's nonempty, there is
at least one point $c \in [a]_{\sim_U} \cap [b]_{\sim_V}$, meaning
that $c \sim_U a$ and $c \sim_V b$. We now want to claim that
\[[a]_{\sim_U} \cap [b]_{\sim_V} = [c]_{\sim_{U \cap V}}\]

First let's show
\[[c]_{\sim_{U \cap V}} \subseteq [a]_{\sim_U} \cap [b]_{\sim_V}  \]
Suppose we have a $d$ with $d \sim_{U \cap V} c$. Therefore
$d \sim_{U} c$ and $d \sim_{V} c$. By transitivity,
$d \sim_{U} a$ and $d \sim_{V} b$, as required.

Now let's show
\[[a]_{\sim_U} \cap [b]_{\sim_V} \subseteq [c]_{\sim_{U \cap V}}\]
Suppose we have $d$ with $d \sim_U a$ and $d \sim_V b$.
By transitivity, $d \sim_U c$ and $d \sim_V c$.
By previous lemma, we are done.
\cqed
\end{proof}

I think I see what topology this gives for $(b)$ above --- and it's notable
that it's not $T_1$. The two `closed points' have the same neighborhoods, since any neighborhood
in the time dimension lets you project from one point forward and back to the other.

Actually, this is making me think (unfortunately!) that there {\em is} an example of an undesired
function representation even back in fibrations over $[0,1]$. I'm not sure what's stopping me from
having basically
\[
\begin{tikzpicture}[baseline=-1.5em,
   dot/.style={circle,fill=black,inner sep=0pt, minimum size=3pt},
   circ/.style={circle,inner sep=0pt,minimum size=4pt,draw,fill=white},
   bigcirc/.style={circle,inner sep=0pt,minimum size=12pt,draw,fill=white}]
\begin{scope}[scale=1]
\draw plot [smooth, tension=0.8] coordinates {(-0.8,0) (-0.5,-.25) (0,-0.5)};
\draw plot [smooth, tension=0.8] coordinates {(-0.8,-1) (-0.5,-0.75) (0,-0.5)};
\draw plot [smooth, tension=0.8] coordinates {(0.8,0) (0.5,-0.25) (0,-0.5)};
\node[bigcirc]  at (0,-0.5) {};
\node[dot]  at (.07,-0.55) {};
\node[dot]  at (-0.07,-0.45) {};
\end{scope}
\end{tikzpicture}
\]
with {\em two} points right at the conjunction. Fibrant transport is still unique, the only
problem is that those two points have exactly the same neighborhoods so the space as a whole
isn't $T_1$.

Is this example also not a sober space? That might be more relevant. I
think it's not; I think the set of those two points $\{a, b\}$ is an
irreducible closed set that is the closure of either $a$ or $b$
independently.
\subsubsection*{Functoriality of $\Pi$ take 2}

Consider a function $\I \to \I$ such as $k = (\dns * \ups\dns) * \ups\ups$.
If I try to coerce along it, like $\rcoe_{i.X(k(i))}(a)$ I should conclude
\[\rcoe_{i.X(k(i))}(a) \equiv \rcoe_{X \o (\dns * \ups\dns) * \ups\ups}(a)\]
\[ \equiv \rcoe_{X \o (\dns * \ups\dns)} \rcoe_{X \o \ups\ups}(a)\]
\[ \equiv \rcoe_{X \o \dns} \rcoe_{X \o \ups\dns} \rcoe_{X \o \ups\ups}(a)\]
\[ \equiv \rcoe_{X \o \dns} \rcoe_{X \o \ups} (a)\]
\[ \equiv \rcoe_{X} (a)\]
So maybe aiming for maximal decomposition of $*$ would work better?
Although the problem is still probably rooted in the connections $i\lor j$ or $i\land j$ or whatever.
Hmm.

\subsubsection*{Sober spaces and frames and such }
How can I think above the fibrancy condition if I'm dually talking about algebras of open sets?
I could draw the diagram
\[
\begin{tikzcd}
  2 \ar[from=r, "x"']\ar[from=d, "0"] & X \ar[from=d]\\
\Omega\I \ar[from=ur, dotted]\ar[from=r, "f"] & \Omega\I
\end{tikzcd}
\]
or upside down so that the cofibration is on the left:
\[
\begin{tikzcd}
 \Omega \I \ar[r, "f"]\ar[d] & \Omega\I \ar[d, "0"]\\
X \ar[ur, dotted]\ar[r, "x"'] & 2
\end{tikzcd}
\]
If $f : \I \to \I$ is monotone continuous, what characterizes $f^{-1}
: \Omega\I \to \Omega \I$? At least it should be the case that it maps
not just open sets to open sets, but interals to intervals. Is that
enough? No, that would also include antitone functions. Something
about how wiggling the bottom coordinate of the interval affects only
the bottom coordinate of the output. I think a roughly appropriate function
$g : \Omega \I \to \Omega \I$ arises from a suitable $h : \I \to \I$
as $g(a,b) = (h(a), h(b))$, where $h$ is mostly injective (except for
at 0 and 1) surjective and monotone (but not necessarily continuous).

\subsubsection*{Giving up on that direction for now}

Trying to think about things dually seems a bit too hard. Let me proceed by
believing that a path in $\rset$ is {\em some} kind of sufficiently well-behaved space
--- probably at least $T_1$ and/or sober, something that excludes the double-point scenario above ---
over $\I$.

What then is a $2$-path in $\rset$? Clearly at least homotopically gradual reparameterizations of $\I_1$
should be allowed. Like --- if $i : \I_1 \prov f, g : \I_1$ are monotone, then there
should be an interpolant
\[i : \I_1, j : \I_2 \prov j[f,g]: \I_1\]
such that
\[i : \I_1 \prov 0[f,g] \equiv f(i)\]
\[i : \I_1 \prov 1[f,g] \equiv g(i)\]
So I expect by mere substitution compatibility that at least
\[ i : \I_1, j : \I_2 \prov X(j[f,g]) : \rset\]
to be valid, if $X : \I_1 \to \rset$ and $f,g : \I_1 \to \I_1$.

\subsubsection*{Yoneda}

Start with the thought
\[\hom(yc, F) = F(c)\]

Imagine that $F : C^\op \to \rset$.
Then it must be that $c : C$. The function $y$ is defined by
\[y : C \to (C^\op \to \rset)\]
\[y\ c\ d = d \le c \]
and of course
\[\hom : C^\op \to C \to \rset\]
\[\hom( c, d) = c \le d \]
So we're really saying
\[(\lambda d . d \le_C c) \le_{C^\op \to \rset} F = F(c)\]
Univalence tells us to try to find an isomorphism. Suppose I have
a path $(\lambda d . d \le_C c) \le_{C^\op \to \rset} F$.

Right away I have a problem. I want to transport from the identity path, but how do I even type it?
Maybe with some clever shifting around? Let me suppose I could do it. Then maybe I could derive
$(c \le_C c) \le_{\rset} F\ c$. Then we put in the identity path and get out $F\ c$.

In the other direction, suppose we have an $F\ c$. Then we want to make a path
$(\lambda d . d \le_C c) \le_{C^\op \to \rset} F$. Maybe it suffices, for every $d: C^\op$, to make a path
$(d \le_C c) \le_{ \rset} F\ d$ which is essentially a function
$(d \le_C c) \to F\ d$. And if transport on $F$ works, then we can transport $F\ c$ back to $F\ d$.

\subsubsection*{Identity Paths}

Let's think about the fact that I can't naively have
\[a : A \prov \ridp : a \le_A a \]
since ostensibly what I'm supposed to have is
\[a : A^\op, b : A \prov a \le_A b : \rset\]
The way I want to weasel out of this is by saying something like
\[a : \square A \prov \ridp : a \le_A a \]
where $\square A$ is some shifted type from which I can make (i.e. is a subtype of) both $A$ and $A^\op$.

Let's try to define this carefully. Either the $\I_1$-paths of $A$ correspond to $\I_2$-paths of $\square A$
or vice versa. Things I know should exist include:
\begin{enumerate}
\item $\I_2 \to \I_1$
\item $\I_2 \to \I_1^\op$
\item $\square A \to A$
\end{enumerate}

Ok, so I should be able to map a path $\I_1 \to \square A$ into
$\I_1 \to A$. Let's assume $\I_1 \to \square A$ is like $\I_2 \to A$. Then this is
asking me to map $\I_2 \to A$ into $\I_1 \to A$. This would require $\I_1 \to \I_2$,
which is backwards from what I have.

So I conclude $\I_2 \to \square A$ is like $\I_1 \to A$. Mapping
$\square A$ into $A$ requires
mapping $\I_2 \to \square A$ into $\I_2 \to A$ which is equivalent to
mapping $\I_1 \to A$ into $\I_2 \to A$, which can be achieved with $\I_2 \to \I_1$, which I have.
I could have also reasoned that I need
$\I_1 \to \square A$ into $\I_1\to A$ which is like $\I_1 \to \square A$ into $\I_2 \to \square A$, which
can also be achieved with $\I_2 \to \I_1$, which I have.

That leaves me not really knowing what the $\I_1$-paths in $\square A$ are. But I think the $\I_{n+1}$ paths
in $\square A$ are the $\I_n$-paths in $A$.

This sort of makes sense. $\square A$ is less connected than $A$ or $A^\op$, so it can map into them.
It might also be that I'm somehow supposed to do the opposite shift, like
\[a :  A \prov \ridp : a \le_{\lx A} a \]
but I don't know how to pick. The approach with $\square$ maybe seems better to the extent that I think I ought to
at least have identity paths for every type, not just laxified types.

What would intro/elim rules look like for $A$?
\[
\erule
{i : \I_n; \cdot \prov e : A}
{i : \I_{n+1}; \cdot \prov \square e : \square A}
\]

\[
\erule
{
{i : \I_{n+1}; \cdot \prov  e : \square A}
\qquad
{ x : A \prov  e' : C}
}
{i : \I_n; \cdot \prov \blet \square x = e \bin e': C}
\]

\subsection*{Fibrations}
Suppose we have $p : \E \to \B$. Let's make up a category $\Sigma p$ whose objects are $\E_0 \cup \B_0$,
and whose homsets are given by:
\[\begin{tabular}{l}
$ \Sigma p [E, E'] = \E[E, E']$\\
$ \Sigma p [E, B] = \B[pE, B]$\\
$ \Sigma p [B, B'] = \B[B, B']$\\
$ \Sigma p [B, E] = \emptyset$
\end{tabular}
\]
So we have all the morphisms from $\B$ and $\E$, and also some `transverse' morphisms $E \to B$.
To compose $f : E \to E'$ and $g : E' \to B'$ we just compose
$pf ; g$.

Say $\Sigma p$ `has tranverse pullbacks' if it has all
pullbacks of transverse morphisms against $\B$-morphisms, and the pullback
of a transverse identity is another transverse identity.
\begin{lemma}
The functor $p$ is a grothendieck fibration iff $\Sigma p$ has transverse pullbacks.
\end{lemma}

\begin{proof}
Suppose the desired pullbacks exist. Then we want to show that $p$ is a fibration.
We need to show that every morphism $f : B' \to B$ has a cartesian lifting for a given
$E$ that lies over $B$. Well, `$E$ lies over $B$' means $pE = B$, so we actually have $f : B' \to pE$,
and the identity $\rid_{pE}$ is in $\Sigma p[E, pE]$,
so we can ask for the pullback
\[\begin{tikzcd}
X \ar[r, dotted, "\phi"]\ar[d,dotted, "\pi"']\pbck & E\ar[d, "\rid_{pE}"] \\
B' \ar[r, "f"'] & pE \\
\end{tikzcd}\]
Now observe that $X$ can't come from $\B$, because $\phi : X \to E$. So let's rename $X$ to $E' \in \E$:
\[\begin{tikzcd}
E' \ar[r,  "\phi"]\ar[d, "\pi"']\pbck & E\ar[d, "\rid_{pE}"] \\
B' \ar[r, "f"'] & pE \\
\end{tikzcd}\]
Anyhow we said we had transverse pullbacks, so what we have is actually
\[\begin{tikzcd}
E' \ar[r,  "\phi"]\ar[d, "\rid_{pE'}"']\pbck & E\ar[d, "\rid_{pE}"] \\
pE' \ar[r, "p\phi"'] & pE \\
\end{tikzcd}\]
Ok, is $\phi$ actually cartesian? We let $E''$ and $\psi : E'' \to E$ and $g : pE'' \to pE$ be given, and the pullback
gives us $\chi$:
\[\begin{tikzcd}
E'' \ar[dr, dotted, "\chi" description] \ar[drr, bend left, "\psi"] \ar[d, "\rid_{pE''}"']\\
p E''\ar[dr, bend right, "g"']&E' \ar[r,  "\phi"]\ar[d, "\rid_{pE'}"']\pbck & E\ar[d, "\rid_{pE}"] \\
&pE' \ar[r, "p\phi"'] & pE \\
\end{tikzcd}\]
%
In the other direction, assume $p$ is a fibration. Let's try to find a pullback of
\[\begin{tikzcd}
X \ar[r, dotted, "\phi"]\ar[d,dotted, "\pi"']\pbck & E\ar[d, "\rho"] \\
B' \ar[r, "f"'] & B \\
\end{tikzcd}\]
We know that $\rho$, being transverse, factors as
\[\begin{tikzcd}
X \ar[r, dotted, "\phi"]\ar[dd,dotted, "\pi"']\pdbck & E\ar[d, "\rid_{pE}"] \\
& pE \ar[d, "\rho_0"]\\
B' \ar[r, "f"'] & B \\
\end{tikzcd}\]
Maybe we need to ask that $\B$ has pullbacks! Then we could do
\[\begin{tikzcd}
X \ar[r, dotted, "\phi"]\ar[d,dotted, "\pi"']\pbck & E\ar[d, "\rid_{pE}"] \\
B'' \pbck \ar[r] & pE \ar[d, "\rho_0"]\\
B' \ar[r, "f"'] & B \\
\end{tikzcd}\]
and ask the fibration oracle for a cartesian $\phi$;
\[\begin{tikzcd}
E' \ar[r, "\phi"]\ar[d, "\rid_{pE'}"']\pbck & E\ar[d, "\rid_{pE}"] \\
pE' \ar[d, "\rho_1"'] \pbck \ar[r] & pE \ar[d, "\rho_0"]\\
B' \ar[r, "f"'] & B \\
\end{tikzcd}\]
Then we could use the cartesianness of $\phi$ to verify that the top square {\em is} in fact a pullback,
and maybe we could at least reason that if $\rho_0$ is an isomorphism, then $\rho_1$ must also be?
\cqed?
\end{proof}

Not totally sure if this proof actually works, or if it does, whether it accomplishes what I want. What
about comparing the requirement that $\Pi p$ has pullbacks to a pseudofunctor $\B^\op \to \rset$?

\subsection*{Pictures of Functions}
[2019.12.20]

If I had a type $A$ which was, more or less, the walking adjunction, i.e.
the category with two objects $\C$ and $\D$, and two nontrivial
morphisms $F : \C \to \D$ and $U : \D \to \C$, and two 2-cells
$\epsilon : FU \to \rid_\D$ and $\eta : \rid_\C \to UF$, satisfying the
triangle equalities in some appropriate way, then I reckon that paths
\[ p, q : \C \le_A \C \]
would be effectively totally ordered sets, in the sense that a path
\[ f : p \le_{\C \le_A \C} q \]
would be an order-preserving map from $p$ to $q$.

For example, a 3-element set would be represented by the path
\[UFUFUF : \C \le_A \C\]
and an order-preserving map that collapses two of those elements would be
\[\rid_U * \epsilon * \rid_{FUF} : UFUFUF \le UFUF\]

Another reasonable goal is expressing ordinary non-ordered sets. I might start
by trying to assert the existence of (internal to a type $B$)
\[\hbox{points }\ \C, \D\]
\[\hbox{paths }\ F : \C \le \D,\quad G: \D \le \C\]
\[\hbox{2-cells }\ \eta : \rid_\C \le UF,\quad \epsilon': UF \le \rid_\C\]
\[\hbox{3-cells }\ H : \rid_{\rid_\C} \le \epsilon'  \eta,\quad M: \eta \epsilon' \le \rid_{UF}\]
and now paths
\[p : \rid_\C \le_{\C \le_B \C} \rid_\C\]
are {\em something} like sets, and for every function $p \to q$ there does exist some way of getting
\[p \le_{\rid_\C \le_{\C \le_B \C} \rid_\C} q\]
However, there are {\em too many} ways, since topologically we keep track of how such functions are {\em braided}.

A question I lack sufficient intuition to answer even intuitively is: if we hop up {\em just one more dimension},
do we get actual sets and functions, or is there some breakdown of the correspondence at still higher dimensions?

What does it look like if I go up one more dimension?
\[\hbox{points }\ P_0, P_1\]
\[\hbox{paths }\ F_0 : P_0 \to P_1,\quad F_1: P_1 \to P_0\]
\[\hbox{2-cells }\ \alpha_0 : \rid_{P_0} \to F_1F_0,\quad \alpha_1: F_1F_0 \to \rid_{P_0}\]
\[\hbox{3-cells }\ m_0 : \rid_{\rid_{P_0}} \to \alpha_1\alpha_0,\quad m_1 : \alpha_1\alpha_0 \to \rid_{\rid_{P_0}} \]
\[\hbox{4-cells }\ \eta : \rid^3_{P_0} \to m_1m_0,\quad \mu : m_0m_1 \to \rid_{\alpha_1\alpha_0} \]
Notice how the two $4$-cells we postulate break the pattern of the previous two dimensions; instead of showing
how a sphere-like shape can appear and disappear, it shows (with $\eta$) how a sphere can {\em appear},
but (with $\mu$) how a pair of hemispheres can be joined.

Here the things that seem like they might behave like sets are the 3-cells
\[ p : \rid^2_{P_0} \to \rid^2_{P_0}\]
We also can say that $\rid^3_{P_0}$ is the empty set, and $m_1 m_0$ is the $1$-element set.
So $\eta$ is simply the unique map $0 \to 1$. Now $\mu$ is {\em not} quite the unique map $2 \to 1$,
but that map (which has type $m_1m_0m_1m_0 \to m_1m_0$) can straightforwardly be derived from
it by precomposition and postcomposition of $m_0$ and $m_1$.

Now what happens if I start with $\rset$ and work backwards? I say
that I think $\rset$ is equivalent to loops in some type $S$, namely
\[\rset = (\iota \to_{S} \iota)\]
For $\iota : S$ that plays the role of $\rid^2_{P_0}$. Here $S$ is like $\rid_{P_0} \to \rid_{P_0}$.
I notice $\alpha_1\alpha_0 : \rid_{P_0} \to \rid_{P_0}$, so it's also an element of $S$.
Call it $\upsilon : S$. The 2-cell $\upsilon$ is the `one-element set' of a two-dimensional slice of space.
I believe in the existence of $m_0 : \iota \to \upsilon$ and $m_1 : \upsilon \to \iota$.
And $\eta : \rid_\iota \to m_1 m_0$ and $\mu : m_0 m_1 \to \rid_{\upsilon}$.

Now $S$ itself should also be a loop space.
\[S = (\iota' \to_{S'} \iota')\]
so the analogy goes
\[
\begin{tabular}{l|l}
$\iota$&$\rid^2_{P_0}$\\
$\iota'$&$\rid_{P_0}$\\
$S$&$\rid_{P_0} \to \rid_{P_0}$\\
$S'$&${P_0} \to {P_0}$\\
$\upsilon$&$\alpha_1\alpha_0$\\
\end{tabular}
\]

Actually, wait, maybe the immediate precursor to $\rset$ has a simple description unto itself.
Let's still call it $S$. It has elements $\upsilon$ and $\iota$, which are `1-element' and `0-element' sets.
Can I to hope for the equivalence of types
\[S = (\iota \to_{S} \iota)\]
? No, actually probably not, because then I can't believe that $\rset$ satisfies that same property. Let's try to imagine a table
like so:
\[ C'' : \iota'' \to \upsilon'',\quad D'': \upsilon'' \to \iota''\]
\[ C' : \iota' \to \upsilon',\quad D' : \upsilon' \to \iota' \]
\[ C : \iota \to \upsilon,\quad \mu : C'D' \to \rid_{\upsilon'} \]
\[\upsilon = D'C'\qquad \iota = \rid_{\iota'}\]
\[\upsilon' = D''C''\qquad \iota' = \rid_{\iota''}\]
\[\upsilon'' = D'''C'''\qquad \iota'' = \rid_{\iota'''}\]
expanding out a bit:
\[
\begin{tabular}{ll}
$ C_2 : \iota_3 \to \upsilon_3$ & $ D_2: \upsilon_3 \to \iota_3 $\\
$ C_1 : \iota_2 \to \upsilon_2$ & $ D_1: \upsilon_2 \to \iota_2 $\\
$ C_0 : \iota_1 \to \upsilon_1$ & $ D_0 : \upsilon_1 \to \iota_1 $\\
$ \eta : \iota_0 \to \upsilon_0$ & $ \mu : C_0D_0 \to \rid_{\upsilon_1} $
\end{tabular}
\]
\[\iota_n = \rid_{\iota_{n+1}} : \iota_{n+1} \to \iota_{n+1}\]
\[\upsilon_n = D_n C_n : \iota_{n+1} \to \iota_{n+1}\]
Ok, so $\eta$ actually is a path in $\rset$, i.e. a function from the 0-element set to the 1-element
set. That means $D_0C_0$ is supposed to be an {\em element} of
$\rset$, but it looks like a path. So let's say it's a path in $S_0$. So $D_0, C_0$ are paths in $S_0$.
And $D_i, C_i$ are paths in $S_i$. An object of $\rset$ is a path $1 \to 1$ in $S_0$.

Ok, so that means $\iota_{n+1}, \upsilon_{n+1}$ are supposed to be {\em elements} of $S_n$. And
\[ S_n = (\iota_{n+2} \to_{S_{n+1}} \iota_{n+2}) \]
Hm, not sure I like this. Let's try reindexing:

\[
\begin{tabular}{ll}
$ C_2 : \iota_2 \to \upsilon_2$ & $ D_2: \upsilon_2 \to \iota_2 $\\
$ C_1 : \iota_1 \to \upsilon_1$ & $ D_1: \upsilon_1 \to \iota_1 $\\
$ C_0 : \iota_0 \to \upsilon_0$ & $ D_0 : \upsilon_0 \to \iota_0 $\\
$ \eta : \rid_{\iota_0} \to D_0C_0$ & $ \mu : C_0D_0 \to \rid_{\upsilon_0} $
\end{tabular}
\]
\[\iota_n = \rid_{\iota_{n+1}} : \iota_{n+1} \to \iota_{n+1}\]
\[\upsilon_n = D_{n+1} C_{n+1} : \iota_{n+1} \to \iota_{n+1}\]
Here $\rset = \iota_0 \to \iota_0$ and $S_n = \iota_{n+1} \to \iota_{n+1}$.

\subsubsection*{Starting Fresh}
Postulate:
\[
\begin{tabular}{ll}
$  \iota_n : S_n $ & $ \upsilon_n : S_n $\\
$ C_n : \iota_n \le \upsilon_n$ & $ D_n: \upsilon_n \le \iota_n $\\
$ \eta : \rid_{\iota_0} \le D_0C_0$ & $ \mu : C_0D_0 \le \rid_{\upsilon_0} $
\end{tabular}
\]
Postulate an equivalence $\phi : S_n \to (\iota_{n+1} \le \iota_{n+1})$, then assert
\[\phi(\iota_n) = \rid_{\iota_{n+1}} : \iota_{n+1} \le \iota_{n+1}\]
\[\phi(\upsilon_n) = D_{n+1} \o C_{n+1} : \iota_{n+1} \le \iota_{n+1}\]


Then I can `read off' $\rset$ as whatever $\iota_0 \le \iota_0$  turns out to be;
the empty set is $\rid_{\iota_0}$, and the singleton is $D_0C_0$, and union is composition.

\subsubsection*{Delooping}
What are the sets and morphisms of the deloopings of $\rset$? Naively they look like sets,
and something like injective relations between sets with some collateral noise $Y:\rset$, i.e. spans
\[
\left(
\begin{tikzcd}
  &X\ar[dl, >->]\ar[dr, >->]\\
A && B
\end{tikzcd}, Y
\right)
\]
where all arms are mono. Actually maybe this is better represented as
$X_\emptyset, X_{A}, X_{B}, X_{A,B}$, which is a `morphism'
from $X_A + X_{A, B}$ to $X_B + X_{A, B}$.

Better to say: A morphism from $A$ to $B$ is some sets
\[(X_\emptyset, X_{A}, X_{B}, X_{A,B})\]
and some isos $A \cong X_A + X_{A, B}$ and $B \cong X_B + X_{A, B}$.
Then the composition of $X : A \to B$ and $Y : B \to C$ is
\[(X_\emptyset + Y_\emptyset + X_B \times_B Y_B, \]
\[X_A + X_{A, B} \x_B Y_B,\]
\[ X_B \x_B Y_{B, C} + Y_C,\]
\[ X_{A,B} \x_B Y_{B,C} ) \]

\subsubsection*{Triangle Equalities}

I was mistaken above thinking just the $DC \leftrightarrow 1$ cells were sufficient.
For a composition of, e.g. an arbitrary function $1 \to 2$ followed by $!: 2\to 1$
to behave correctly, we do have to have triangle equalities for the adjunction.

So I start with
\[\begin{tabular}{lll}
  {objects}& $z, u $\\
  {morphisms}& $F : z \to u$& $U : u \to z $\\
  {2-cells}& $\eta : \rid_z \to UF$& $\epsilon : FU \to \rid_u $\\
  {3-cells}& $\dash : \rid_{U} \leftrightarrow (\rid_{U} * \epsilon) \o (\eta * \rid_{U})$\\
  & $\dash : \rid_{F} \leftrightarrow (\epsilon * \rid_{F}) \o (\rid_{F} * \eta)$
\end{tabular}\]

\subsubsection*{Homotopy Types as Spaces Concretely}

If I have all four of $\eta : 1 \leftrightarrow UF : \eta'$ and $\epsilon : FU \leftrightarrow 1 : \epsilon'$, then I can make
arbitrary non-simply-connected 2d shapes as well. E.g. here's a circle type:
\[ \eta' \o (\rid_U * (\epsilon \o \epsilon') * \rid_F) \o \eta\]
I speculate that some appropriate set
of higher dimensional cells that is {\em not} the fully ramified tree of equivalence cells
would express what kind of {\em functions} could be defined, which behave like HoTT functions
with respect to holes in spaces.
\end{document}
