\documentclass{article}
\usepackage{amssymb}
\input{theorem}

\input{prooftree}
\def\erule#1#2{\begin{prooftree}#1\justifies #2\end{prooftree}}
\def\pair#1#2{\langle #1 , #2 \rangle}

\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{calc}
\definecolor{morange}{rgb}{1,0.56,0}
\definecolor{lorange}{rgb}{1,0.95,0.8}
\definecolor{mgreen}{rgb}{0,0.56,0}
\definecolor{lgreen}{rgb}{0.95,1,0.8}
\definecolor{mblue2}{rgb}{0,0.2,1.0}
\definecolor{lblue}{rgb}{0.8,0.95,1}
\definecolor{mred}{rgb}{0.9,0.1,0.1}
\definecolor{mgreen}{rgb}{0.1,0.5,0.1}
\definecolor{mblue}{rgb}{0.3,0.3,0.9}
\def\bitf#1{#1 [smooth, tension=0.8] coordinates {(-1.6,2) (-1,1) (0,0)}}
\def\bitg#1{#1 [smooth, tension=0.8] coordinates {(1,2) (0.55,1) (0,0)}}
\def\bitgrev#1{#1 [smooth, tension=0.8] coordinates {(0,0) (0.55,1) (1,2)}}
\def\bitfg#1{#1 [smooth, tension=0.8] coordinates {(0,0) (0,-2) }}
\def\ep{\varepsilon}
\def\bcase{\mathop\mathbf{case}}
\def\bof{\mathop\mathbf{of}}
\def\binj{\mathbf{inj}}
\def\blet{\mathrel\mathbf{let}}
\def\bin{\mathrel\mathbf{in}}
\def\bmatch{\mathrel\mathbf{match}}
\def\bwith{\mathrel\mathbf{with}}
\def\pbck{\ar[dr, phantom, pos=0, "\lrcorner"]}
\def\ups{{\uparrow}}
\def\dns{{\downarrow}}
\def\grad{\nabla}
\def\uups{{\Uparrow}}
\def\ddns{{\Downarrow}}
\def\adjust{\big|}
\def\O{\mathcal{O}}
\def\rid{\mathsf{id}}
\def\ridp{\mathsf{idp}}
\def\rcoe{\mathsf{coe}}
\def\rtype{\mathsf{type}}
%\def\int{\square}
\def\bd{\partial}
\def\prov{\vdash}
\def\prequiv{\dashv\vdash}
\def\imp{\Rightarrow}
\def\cqed{\hskip2mm{\vrule width .5em height .5em depth 0em}} % at the end of a |P.
\def\o{\circ}
\def\lx{\bigcirc}
\def\B{\mathbb{B}}
\def\C{\mathbf{C}}
\def\R{\mathbb{R}}
\def\bx{\mathbf{x}}
\def\by{\mathbf{y}}
\def\bv{\mathbf{v}}
\def\bt{\mathbf{t}}
\def\bu{\mathbf{u}}
\def\bp{\mathbf{p}}
\def\bz{\mathbf{0}}
\def\bzee{\mathbf{z}}
\def\S{\mathfrak{S}}
\def\M{\mathbb{M}}
\def\X{\mathbf{X}}
\def\Y{\mathcal{Y}}
\def\x{\times}
\def\st{\mathrel|}
\def\rset{\mathbf{Set}}
\def\rcat{\mathbf{Cat}}
\def\op{\mathsf{op}}
\def\P{\mathbb{P}}
\def\I{\mathbb{I}}
\def\U{\mathbb{U}}
\def\N{\mathbb{N}}
\def\Z{\mathbb{Z}}
\def\tw{\mathbf{2}}
\def\dash{\hbox{---}}
\def\dom{\mathop{\mathrm{dom}}}
\def\cod{\mathop{\mathrm{cod}}}
\def\celse{\mathrel{|}}
\def\cn{{:}}
\def\rok{\mathrel\mathsf{ok}}
\def\llam#1{\langle {#1} \rangle}
\def\hf{{\odot}}
\def\lerp#1#2#3{#1[\![#2,#3]\!]}
\def\nope{\emptyset}
\def\ep{\varepsilon}
\def\rng{\mathop\mathrm{rng}}
\def\npart{\not\!\partial}
\def\null{\emptyset}

\begin{document}
\tikzset{>=stealth}
\tikzset{
   commutative diagrams/.cd,
   arrow style=tikz,
   diagrams={>=stealth}}

\section{Main}

Consider a category $\I$ whose objects are closed intervals in the
reals, and whose morphisms are monotone continuous maps.

We attempt to define a Grothendieck topology on this category by
saying that a sieve $S$ on an interval $X = [a,b]$ is covering iff the union of the ranges
of the maps in $S$ is the whole interval $[a,b]$.

\[ \hom(\dash, X) \supseteq S  \in J(X) \iff X = \bigcup_{g : Z \to X \in S} \rng g \]
\[  \iff \forall x \in X . \exists Z . \exists g : Z \to X \in S . \exists z \in Z . g(z) = x\]

To check this is a Grothendieck topology, we check:

\begin{enumerate}
\item ({\bf Change of Base}) If $S \in J(X)$, and $f : Y \to X$, then $f^* S \in J(Y)$.
Let $S \in J(X)$ and $f : Y \to X$ be given. We know $f^* S = \{ g : Z \to Y \st f \o g \in S\}$.
So we want to show
\[f^*S \in J(Y)\]
\[  \equiv \forall y \in Y . \exists Z . \exists g : Z \to Y \in f^* S . \exists z \in Z . g(z) = y\]
$$  \equiv \forall y \in Y . \exists W . \exists h : W \to Y . fh \in S . \exists w \in W . h(w) = y  \eqno(1) $$
and what we do know is
$$  \forall x \in X . \exists Z . \exists g : Z \to X \in S . \exists z \in Z . g(z) = x \eqno(2) $$
So let $y \in Y$ be given. Set $x$ in $(2)$ to $f(y)$, and let $Z$ and $g: Z \to X$ and $z$ be given.
We now know $g(z) = x$. In $(1)$, set $W$ to the pullback $Y \times_X Z$ of $f$ and $g$, and set $h = f^* g$.

\[\begin{tikzcd}
  Y \times_X Z \ar[r] \ar[d, "h"'] \pbck& Z \ar[d, "g"]\\
  Y  \ar[r, "f"'] & X \\
\end{tikzcd}\]
\end{enumerate}

\section{Oriented Cubes}

\def\inc{\triangleleft}
A cell in a cube is a partial substitution $\Gamma \prov \sigma$ of binary digits for some of the variables in $\Gamma$.
For example, a vertex of a cube is $x,y,z\prov[0/x][1/y][0/z]$ and a face of a cube is $x,y,z\prov [1/y]$. Lower
dimensional cells have more substitutions in them. A lower-dimensional cell is incident to a one-dimension-higher cell
if it contains one more substitution: the relation $\sigma \inc \rho$  is defined by
\[
\erule
{x\in \Gamma}
{\Gamma \prov \sigma[b/x] \inc \sigma }
\]

We aim to define two more relations, $\sigma \dns \rho$ (resp. $\sigma \ups \rho$), which indicate that $\sigma$
is in the domain (resp. codomain) of $\rho$. These two relations partition $\inc$: every face incident to a one-higher cell
is either in its domain or codomain.

We can index all three of these relations by $n$, the number of variables in the context.

\def\s{{*}}
\subsection{Maybe Not Substitutions}

The more concrete way of looking at this is with strings of $\{0,1,\s\}$. The vertices of the edge $\s$ are $0,1$.
The vertices of the face $\s\s$ are $00,01,10,11$. The edges are $0\s,1\s,\s0,\s1$. The domain and codomain relations $\dns$
and $\ups$ operate on pairs of strings of the same length such that the second one has one more $\s$.
We have $0 \dns \s$ and $1 \ups \s$. I think I want to assert
\[0\s \dns \s\s \qquad \s1 \dns \s\s \qquad \s0 \ups \s\s \qquad 1\s \ups \s\s \]
The main principle I want to impose goes like this. Let's say $f$ is a function on strings that inserts a binary
digit in a fixed place in a string. Then
\[
\erule
{s \dns t}
{f(s) \dns f(t)}
\qquad
\erule
{s \ups t}
{f(s) \ups f(t)}
\]
Now let's look at the cube $\s\s\s$. By the above principle, I infer all of
\[00\s \dns 0\s\s \qquad 0\s1 \dns 0\s\s \qquad 0\s0 \ups 0\s\s \qquad 01\s \ups 0\s\s \]
\[00\s \dns \s 0\s \qquad \s 01 \dns \s 0\s \qquad \s 00 \ups \s 0\s \qquad 10\s \ups \s 0\s \]
\[0\s 0\dns \s \s 0 \qquad \s10 \dns \s\s 0 \qquad \s00 \ups \s\s 0 \qquad 1\s 0\ups \s\s 0 \]
\[10\s \dns 1\s\s \qquad 1\s1 \dns 1\s\s \qquad 1\s0 \ups 1\s\s \qquad 11\s \ups 1\s\s \]
\[01\s \dns \s 1\s \qquad \s 11 \dns \s 1\s \qquad \s 10 \ups \s 1\s \qquad 11\s \ups \s 1\s \]
\[0\s 1\dns \s \s 1 \qquad \s11 \dns \s\s 1 \qquad \s01 \ups \s\s 1 \qquad 1\s 1\ups \s\s 1 \]

I notice that there are some cells that are {\em balanced} in the sense that they occur as a domain and a codomain,
and some that are {\em imbalanced}, occurring twice as domain or twice as codomain. The imbalanced (twice-domain) are
\[00\s \qquad 0\s 1 \qquad \s 11\]
The imbalanced (twice-codomain) are
\[11\s \qquad 1\s 0 \qquad \s 00\]
and the balanced are
\[10\s \qquad 01\s \qquad \s 01 \qquad \s 10 \qquad 0\s 0 \qquad 1\s 1\]

So there should be a unique partition of the faces of the cube such
that each half covers the domain-imbalanced edges once as domains, and
covers the codomain-imbalanced edges once as codomains, and covers
every other edge once as domain and once as codomain. This will also induce a partition of the balanced edges.

Let's start with the face $0\s\s$. This forces the edges $00\s$, (imbalanced)
 $0\s1\dns$ (also imbalanced) and $0\s0\ups$ (balanced) and
$01\s\ups$ (balanced). So we have to balance with faces $\s\s0$ and $\s1\s$. This should be sufficient; we end up with

\[00\s \dns 0\s\s \qquad 0\s1 \dns 0\s\s \qquad 0\s0 \ups 0\s\s \qquad 01\s \ups 0\s\s \]
\[0\s 0\dns \s \s 0 \qquad \s10 \dns \s\s 0 \qquad \s00 \ups \s\s 0 \qquad 1\s 0\ups \s\s 0 \]
\[01\s \dns \s 1\s \qquad \s 11 \dns \s 1\s \qquad \s 10 \ups \s 1\s \qquad 11\s \ups \s 1\s \]

Ok, so the two halves of the cube are $\{0\s\s,\s1\s,\s\s0\}$ and $\{1\s\s,\s0\s,\s\s1\}$. Can I convince myself the two
halves of the hypercube are
\[\dom = \{0\s\s\s,\s1\s\s,\s\s0\s,\s\s\s1\} , \cod = \{1\s\s\s,\s0\s\s,\s\s1\s,\s\s\s0\}\]
? Yeah, that seems right.

If so, I can conjecture a rule:
\begin{enumerate}
\item
  $s$ is in the domain of $t$ if $s$ arises from replacing the $n^{th}$ $\s$ in $t$ with $n \mathop\% 2$.
\item
  $s$ is in the codomain of $t$ if $s$ arises from replacing the $n^{th}$ $\s$ in $t$ with $(n+1) \mathop\% 2$.
\end{enumerate}

The `intersection' of any two cells $a,b$ within the domain (or within
the codomain) will be balanced --- in the domain of one of $a,b$ and
the codomain of the other. The `intersection` of a domain cell $d$ and a codomain cell $c$ will either
be in the domain of both $c,d$ or in the codomain of both $c,d$.

Ah, here's a simple definition: we define an orientation  relation $x \sim_b y$ where $x,y\in \{0,1,\s\}^*$.
Let $b,c$ range over $\{0,1\}$.
We assert rules:
\[
\erule
{}
{bx \sim_b \s x}
\qquad
\erule
{x \sim_b y}
{c x \sim_b c y}
\qquad
\erule
{x \sim_b y}
{\s x \sim_{\bar b} \s y}
\]

\subsection{Categorical Description}
\vfil\eject
\def\repl#1#2{[#1/\s_{#2}]}
Fix an $n$. We describe a higher-dimensional category that takes the shape of the $n$-dimensional cube.
Its objects will be binary strings that live in $\{0,1\}^n$. To describe the morphisms and higher cells we do the following.

Let $P_m$ be the set of strings over $\{0,1,\s\}^n$ such that exactly $m$ characters
in the string are $\s$. Obviously this partitions the set of strings of length $n$:
\[ \bigcup_{0 \le m \le n} P_m = \{0,1,\s\}^n \]
The set $P_m$ will be the `primitive' cells of dimension $m$, such that all other cells are compositions of them.
To define which compositions are valid we need to define directed cell boundaries.
When $b \in \{0,1\}$, and $s \in P_m$, and $0 \le i < m$, let $\repl b i s$ be the result of
replacing the $i^{th}$ occurrence (counting zero-indexed) of $\s$ in $s$ with the bit $b$.

We can define the {\em boundary} of a cell of dimension $m$,
which lives in $\Z[P_{m-1}]$.
If $c \in P_m$, we say
\[ \partial c = \sum_{i = 0}^m (-1)^i(\repl 1 i c - \repl 0 i c)\]
For example, $\partial (\s 1 \s) = 11\s - 01\s - \s11 + \s10$. This notion of boundary extends in an evident
way to subset $X \subseteq P_m$: we say
\[ \partial X = \sum_{x \in X} \partial x\]

Now we can define $C_m$, the set of cells of dimension $m$. We say $C_0 = P_0 = \{0,1\}^n$.
For higher dimensions, if $a, b \in C_m$, we say that an element of the homset $C_{m+1}[a, b]$ is
a subset $c \subseteq P_m$ such that $b - a = \partial c$.

Composition is always given by unions of sets, and identities morphisms are always realized by empty sets.
This means that associativity, interchange laws, and unit laws, come for free. What we need to verify
is that compositions even {\em exist} in the right homset,
which means checking the boundary conditions. But this is fairly easy.

To check vertical composition let two cells $\alpha \in C_m[x, y]$ and $\beta \in C_m[y,z]$ be given.
We know therefore that $\partial \alpha = y - x$ and $\partial \beta = z - y$. Therefore
\[\partial(\beta \o \alpha) = \partial(\beta \cup \alpha) = \partial\beta + \partial \alpha = z - y + y - x = z - x\]
so the composition is indeed a cell in $C_m[x,z]$.

To check horizontal composition let two cells $\alpha \in C_m[x_1, y_1]$ and $\beta \in C_m[x_2,y_2]$ be given,
so that $x_1,x_2$ are composable and $y_1,y_2$ are composable. We want to check that $\beta * \alpha$ is a cell in
$ C_m[x_2 \o x_1, y_2 \o y_1]$. So we need
\[ \partial(\beta * \alpha) = y_2 \o y_1 - x_2 \o x_1 \]
but
\[ \partial(\beta * \alpha) = \partial(\beta \cup \alpha) = \partial \beta + \partial \alpha  = y_2 - x_2 + y_1 - x_1\]
so we're done.

\subsection{Domains and Codomains}

If I take the square $\s\s$ and take the set of all its edges $\{0\s, 1\s, \s0, \s1\}$, and take its boundary,
I get $2 \cdot (11 - 00)$. In general if I take the boundary of the $n-1$-cells of the $n$-cube, I expect to get
two times some difference, and I can iterate on taking boundaries of those nonprimitive cells. Let's try to compute
the boundary of all faces in the cube. I think it is twice
\[(11\s + 1\s0 + \s00) - (00\s + 0\s1 + \s11) \]
Because look at $11\s$ in particular, and consider which faces it could have come from, and with which signs.
We have a positive amplitude $\s 1 \s \to 11 \s$ and positive $1\s1 \to 11\s$, because in both those situations it's
the $0^{th}$ $\s$ becoming a 1.

What's the boundary of all the volumes in the hypercube? I think it's
\[(11\s\s + 1\s0\s + 1\s\s1 + \s00\s + \s0\s1 + \s\s11)  -\]
\[(00\s\s + 0\s1\s + 0\s\s0 + \s11\s + \s1\s0 + \s\s00)  \]
In general I think it's
\[ \sum_{i \ne j \atop 0 \le i,j < n} \repl {A_{i}} i \repl {A_{j-i-1}} j \s^n  - \repl {\bar A_{i}} i \repl {\bar A_{j-i-1}} j \s^n\]
where $A_i = 1$ if $i \equiv 0 \pmod 2$ and $0$ otherwise.

\subsubsection{Second Boundaries}

What's the boundary of $(11\s + 1\s0 + \s00)$? It's $111 - 000$.

What's the boundary of $(11\s\s + 1\s0\s + 1\s\s1 + \s00\s + \s0\s1 + \s\s11)$?
It's
\[ (111 \s + 110\s + 11\s1 + 100\s + 10\s1 + 1\s11) \]
\[ {} - (11\s1 + 1\s01 + 1\s11 + \s001 + \s011 + \s111) \]
\[ {} - (110\s + 100\s + 10\s1 + 000\s + 00\s1 + 0\s11) \]
\[ {} + (11\s0 + 1\s00 + 1\s01 + \s000 + \s001 + \s011) \]
%
\[  = (111 \s + 11\s0 + 0\s11 + \s000)    -  (000 \s + 00\s1 + 1\s00 + \s111)\]

\subsubsection{Generalizing}

Let's say $P_m^n$ is the set of strings with length $n$ with exactly
$m$ $\s$s in them. What I'm interested in computing is $\partial^k P^n_{n-1}$.
I notice a pattern that
\[\partial P^2_1 = 2 \cdot (11 - 00)\]
\[\partial^2 P^3_2 = 2^2 \cdot (111 - 000)\]
\[\partial^3 P^4_3 = 2^3 \cdot (1111 - 0000)\]
so I conjecture
\[\partial^m P^{m+1}_m = 2^m \cdot (1^{m+1} - 0^{m+1})\]
If I write $\npart$ for $\partial/2$ then
\[\npart^{m-1} P^{m}_{m-1} = 1^{m} - 0^{m}\]
I think I can define a `star-insertion' operator $S$ such that
\[ S(b_1\cdots b_n) = (b_1\cdots b_n\s) +( b_1\cdots b_{n-1}\s \bar b_n) + \cdots  + (\s \bar b_1 \cdots \bar b_n) \]
for example
\[ S(000) = 000\s + 00\s1 + 0\s 11 + \s 111\]
In which case I conjecture
\[\npart^{k-1} P^{m}_{m-1} = S^{m -  k}(1^{k} - 0^{k})\]





\subsection{Somewhat more Careful Definitions}

Here's an inductive definition of $\partial$, writing $x$ for an arbitrary ternary string in
$\{0,1,\s\}^*$, writing
$b \in \{0,1\}$ and concatenation as $\cdot$ or juxtaposition, and $\epsilon$ for the empty string,
and $\null$ as the unit of the formal sum in $\Z[P^n_m]$. We can lift concatenation from an operation
$P^{n_1}_{m_1} \x P^{n_2}_{m_2} \to P^{n_1+n_2}_{m_1+m_2}$ to
an operation
\[\Z[P^{n_1}_{m_1}] \x \Z[P^{n_2}_{m_2}] \to \Z[P^{n_1+n_2}_{m_1+m_2}]\]
 in the evident linear way.


\[ \partial (\s  x) = 1x - 0x - \s \cdot \partial x\]
\[ \partial (b  x) = b \cdot \partial x\]
\[ \partial \epsilon = \null\]

For example, $\partial 101101 = \null$ and
\[\partial 101\s0\s = 10110\s - 10100\s - 101\s01 + 101\s00\]

Now let's try to define star-insertion inductively from the right.

\[S(x \s) = (S x)\s \]
\[S(x b) = x b \s + (S x) \bar b \]
\[S\epsilon = \s \]

\subsection{Realizing a Big Dumb Mistake}

I'm not keeping track of signs correctly when I compute $\partial^k$. I actually get
$\partial^2 = 0$ as I should have expected. However, I can give a name to the domains and codomains
at each dimension and assert a recurrence relation among them.

Let's define $C^n_m, D^n_m \subseteq P^n_m$. I think it's just going to be as simple as
\[C^n_m = S^m (1^{n-m}) \qquad D^n_m = S^m (0^{n-m}) \]
So for example $D^n_0 = 0^n$ and $C^n_0 = 1^n$ as expected. At the other extreme I have the possibly
slightly weird $D^n_n = C^n_n = \s^n$. Stepping back one step, I have
\[C^n_{n-1} = (1\s\cdots\s) + (\s 0\s\cdots\s) + (\s\cdots \s \pi_{n}) \]
\[D^n_{n-1} = (0\s\cdots\s) + (\s1\s\cdots\s) + (\s\cdots \s \bar \pi_{n}) \]
where $\pi_m$ is the parity of $m$. I think that this satisfies the globular conditions
\[ \partial C^n_m = \partial D^n_m = C^n_{m-1} - D^n_{m-1} \]
for $0 < m \le n$. Let's see which if these I can prove. Let's try $m = 1$. then I'm claiming
\[ \partial C^n_1 = \partial D^n_1 = C^n_{0} - D^n_{0} = 1^n - 0^n \]

Let's try to prove this.
\begin{lemma}
$\partial S(1^{n-1}) = 1^n - 0^n$.
\end{lemma}

\begin{proof}
We know that $S(1^{n-1})$ is equal to
\[\sum_{i = 0}^{n-1} 1^i \cdot \s \cdot 0^{n-1-i}\]
so $\partial S(1^{n-1})$ is
\[\repl 1 0 (\sum_{i = 0}^{n-1} 1^i \cdot \s \cdot 0^{n-1-i}) - \repl 0 0 (\sum_{i = 0}^{n-1} 1^i \cdot \s \cdot 0^{n-1-i})\]
\[=  (\sum_{i = 0}^{n-1} 1^i \cdot 1 \cdot 0^{n-1-i}) -  (\sum_{i = 0}^{n-1} 1^i \cdot 0 \cdot 0^{n-1-i})\]
\[=  (\sum_{i = 0}^{n-1} 1^{i+1}  \cdot 0^{n-1-i}) -  (\sum_{i = 0}^{n-1} 1^i  \cdot 0^{n-i})\]
\[=  (\sum_{i = 1}^{n} 1^{i}  \cdot 0^{n-i}) -  (\sum_{i = 0}^{n-1} 1^i  \cdot 0^{n-i})\]
\[=  1^n + (\sum_{i = 1}^{n-1} 1^{i}  \cdot 0^{n-i}) -  (\sum_{i = 1}^{n-1} 1^i  \cdot 0^{n-i}) - 0^n\]
\[= 1^n - 0^n\]
\cqed
\end{proof}

\begin{lemma}
$\partial S(0^{n-1}) = 1^n - 0^n$.
\end{lemma}

\begin{proof}
We know that $S(0^{n-1})$ is equal to
\[\sum_{i = 0}^{n-1} 0^i \cdot \s \cdot 1^{n-1-i}\]
so $\partial S(0^{n-1})$ is
\[\repl 1 0 (\sum_{i = 0}^{n-1} 0^i \cdot \s \cdot 1^{n-1-i}) - \repl 0 0 (\sum_{i = 0}^{n-1} 0^i \cdot \s \cdot 1^{n-1-i})\]
\[=  (\sum_{i = 0}^{n-1} 0^i \cdot 1 \cdot 1^{n-1-i}) -  (\sum_{i = 0}^{n-1} 0^i \cdot 0 \cdot 1^{n-1-i})\]
\[=  (\sum_{i = 0}^{n-1} 0^{i}  \cdot 1^{n-i}) -  (\sum_{i = 0}^{n-1} 0^{i+1}  \cdot 1^{n-1-i})\]
\[=  (\sum_{i = 0}^{n-1} 0^{i}  \cdot 1^{n-i}) -  (\sum_{i = 1}^{n} 0^{i}  \cdot 1^{n-i})\]
\[=  1^n + (\sum_{i = 1}^{n-1} 0^{i}  \cdot 1^{n-i}) -  (\sum_{i = 1}^{n-1} 0^{i}  \cdot 1^{n-i}) - 0^n\]
\[= 1^n - 0^n\]
\cqed
\end{proof}

Ok, let's now try to think through how we'd prove
\[ \partial C^n_{m+1} = \partial D^n_{m+1} = C^n_m - D^n_m \]
for $0 < m < n$, assuming some reasonable induction hypothesis.
This is the same as showing
\[ \partial S^{m+1} 1^{n-m-1} = \partial S^{m+1} 0^{n-m-1} = S^m 1^{n-m} - S^m 0^{n-m} \]
and we do already know
\[ \partial S^{m} 1^{n-m} = \partial S^{m} 0^{n-m} = S^{m-1} 1^{n-m+1} - S^{m-1} 0^{n-m+1} \]
So let's try to compute
\[\partial S^{m+1} 1^{n-m-1} = \partial S^{m} S 1^{n-m-1}\]
\[= \partial S^{m}( S 1^{n-m-1})\]
\[= \partial S^{m}(  1^{n-m-1}  \s + (S 1^{n-m-2}) 0)\]

and already I'm stuck thinking I need to generalize the i.h. somehow.
Actually let's just try focusing on the i.h. as
\[ \partial S^{m} 1^{k} = \partial S^{m} 0^{k} = S^{m-1} 1^{k+1} - S^{m-1} 0^{k+1} \]
and consider: are there other strings $x$ such that I can say something intelligent about
\[ \partial S^m x\]
?

Let me try to find the sequence $\partial S^m 01$, for example. First we find
\[S(01) = 01\s + 0\s0 + \s10\]
then
\[\partial S(01) = 011 + 010 + 110 - (010 + 000 + 010)\]
\[ = 011  + 110  - (000 + 010)\]
Hmm, a pattern is not yet evident. How about $\partial S 000111$. First we find
\[S(000111) = 000111\s + 00011\s0 + 0001\s00 +   {}\]
\[  000\s000 + 00\s1000 +  0\s11000 +  \s111000 \]
then
\[\partial S(000111) = 0001111 + 0001110 + 0001100 +   {}\]
\[  0001000 + 0011000 +  0111000 +  1111000 \]
\[ -( 0001110 + 0001100 + 0001000 +   {}\]
\[  0000000 + 0001000 +  0011000 +  0111000) \]

\[ = 0001111  +   0111000 +  1111000 \]
\[ -( 0001110 +    0000000 + 0001000    ) \]
Hmm, not a lot clear there either. Here's a wild conjecture:
\[ \partial S^{m} x  = S^{m-1} (\partial S x) \]
For which it would be enough to show
\[ \partial SS x  = S (\partial S x) \]
which is not quite as strong as $S$ and $\partial$ commuting, which doesn't even make sense if $x$ has
no $\s$s in it in the first place.

Let's just see if I can refute or validate this quickly. Let $x$ be the empty string. Then $SSx = \s\s$. The boundary
of that is $1\s - 0\s - \s1 + \s0$. But also $Sx = \s$. And $\partial S x = 1 - 0$. But then
\[ S \partial S =  1\s - 0\s - \s1 + \s0\]
as required. Ok, let $x$ be the string $01$ perhaps. Then $Sx = 01\s + 0\s0 + \s10$, and as we calculated above,
$\partial S (01) = 011 + 110 - (000 + 010)$. If we insert stars again, we find
\[ S\partial S (01) =  011\s + 110\s - (000\s + 010\s) + \]
\[ 01\s0 + 11\s1 - (00\s1 + 01\s1) + \]
\[ 0\s00 + 1\s01 - (0\s11 + 0\s01) + \]
\[ \s100 + \s001 - (\s111 + \s101)  \]
Meanwhile $SSx = 01\s\s + 0\s0\s + \s10\s + 0\s0\s + 0\s\s1 + \s1\s1 + \s10\s + \s1\s1 + \s\s01$, and the boundary of this
is $\partial SSx = $
\[ 011\s + 010\s + 110\s + 010\s + 01\s1 + 11\s1 + 110\s + 11\s1 + 1\s01  \]
\[+  01\s0 + 0\s00 + \s100 + 0\s00 + 0\s01 + \s101 + \s100 + \s101 + \s001  \]
\[- ( 010\s + 000\s + 010\s + 000\s + 00\s1 + 01\s1 + 010\s + 01\s1 + 0\s01) \]
\[- ( 01\s1 + 0\s01 + \s101 + 0\s01 + 0\s11 + \s111 + \s101 + \s111 + \s101 ) \]
and sadly these don't quite look equal. Is this because I messed up some calculation? Let me try on a smaller example.

Consider $x = 0$. Then $Sx = 0 \s + \s 1$. And $SSx = 0\s\s + \s 1 \s + \s 1 \s + \s \s 0 $ and
$\partial SS x = $
\[01\s + 1 1 \s + 1 1 \s + 1 \s 0 \]
\[- (00\s + 0 1 \s + 0 1 \s + 0 \s 0 )\]
\[+ (0\s0 + \s 1 0 + \s 1 0 + \s 0 0) \]
\[- (0\s1 + \s 1 1 + \s 1 1 + \s 1 0) \]

\[=  1 1 \s + 1 1 \s + 1 \s 0 + \s 1 0 + \s 0 0\]
\[- (00\s +  0 1 \s + 0\s1 + \s 1 1 + \s 1 1 )\]
On the other hand $\partial S x  = 01 + 11 - 00 - 01 = 11 - 00$. And so $S\partial S x$ has only 6 terms, not 10.
So sadly I think this conjecture that $\partial SS x = S \partial S x$ is just wrong.

But wait --- what's the difference between them? $S \partial S x = $
\[ 11\s + 1\s 0 + \s 00 - 00 \s - 0\s1 - \s 11\]
So I see
\[ (\partial SS - S \partial S)x = 11\s + \s 1 0 - 01 \s - \s 11\]
Not sure what to do with this.

\subsection{Wait I haven't defined $S^m$ correctly}
When I say
\[ \partial S^{m} 1^{k} = \partial S^{m} 0^{k} = S^{m-1} 1^{k+1} - S^{m-1} 0^{k+1} \]
I actually mean to insert $m$ stars {\em all at once}, not according to the inductive definition I gave above.
I should have that $S^2 0 = 0\s\s + \s 1 \s + \s\s 0$, not
$S^2 0 = 0\s\s + \s 1 \s + \s 1 \s + \s\s 0$. Ok, what happens if I actually keep track of ordering of $\s$s no matter
what, and take a factor of $1\over {m!}$ or something out the very end.

So I define instead $S^n_m : P^n_{m} \to P^{n+1}_{m+1}$ by
\[S^n_m(x b) = x b \s + (S^{n-1}_{m} x) \cdot \bar b \]
\[S^n_m(x \s) = x \s \s + (S^{n-1}_{m-1} x) \cdot \s \]
\[S^0_0\epsilon = \s \]

Now I compute $SS0 = S(0 \s + \s 1) = (0\s\s + 0\s\s + \s 1 \s + \s 1 \s + \s\s0 + \s\s0)$
So let's say $\sigma^m$ is hitting something $m$ times with with $S$ and then scaling by $1 \over m!$.
The conjecture I expect to be true then is
\[ \partial \sigma^{m} 1^{k} = \partial \sigma^{m} 0^{k} = \sigma^{m-1} 1^{k+1} - \sigma^{m-1} 0^{k+1} \]

And now let me reexamine what $\partial \sigma^2 0$ looks like.
We have already $\sigma^2 0 = 0\s\s + \s 1 \s + \s \s 0 $ so
$\partial \sigma^2 x = $
\[01\s + 11\s  + 1\s0\]
\[- (00\s + 01\s + 0\s0)\]
\[ + (0\s0 + \s10 + \s00) \]
\[- (0\s1 + \s11 + \s10) \]
\[=  11\s  + 1\s0 + \s00- (00\s  + 0\s1 + \s11 ) \]
which is the same as $\sigma \partial \sigma x$ now!

Ok, now I have a different hunch: that $S \partial = \partial S$ up to a multiplicative factor
which is the number of stars in the input. Specifically suppose $\partial^n_m : P^n_m \to P^n_{m-1}$
and $S^n_m : P^n_m \to P^{n+1}_{n+1}$. Then I conjecture

\[m \cdot S^n_{m-1} \o  \partial^n_m =  \partial^{n+1}_{m+1} \o  S^n_m : P^n_m \to P^{n+1}_m\]
Where I conventionally say that $\partial^n_0$ always returns $\null$. In particular I want to see
\[ \partial \sigma^2 0 = S \partial S 0\]
so
\[ \partial SS 0 / 2 = S \partial S 0\]
so
\[ \partial S S0  = 2 \cdot S \partial S 0\]
Something seems slightly off here. Let's look at $S 0$. It's $0\s + \s1$. Actually let's consider just $\s$ first.
What's $\partial S \s$? It's $2 (1\s - 0\s - \s1 + \s0)$. What's $S \partial \s$? It's $(1\s - 0\s - \s1 + \s0)$.

Can we do this with a general one-star string? What's $\partial S (x\s y)$? Well,
\[S(x\s y) = Sx \cdot \s \bar y + x \s \cdot S y\]
so
\[\partial S(x\s y) = \partial Sx \cdot \s \bar y + (-1)^{|x|+1}(Sx \cdot 1 \bar y - Sx \cdot 0 \bar y)\]
\[ + (x 1 \cdot S y -  x 0 \cdot S y)  + (-1)^{|x| + 1} x \s \cdot \partial S y\]
while
\[\partial (x \s y) = x 1 y - x 0 y\]
so
\[S \partial (x \s y) = x 1 \cdot S y - x 0 \cdot S y\]
\[{} + S x \cdot 0 \bar y - S x\cdot  1 \bar y\]

Let's try something concrete. What's $S(1\s)$? It's $\s 0 \s + 1\s\s + 1\s\s$.
What's $\partial S (1 \s)$? It's $1 0 \s - 0 0 \s + \s 0 0 - \s 0 1 + 2(11\s - 10\s + 1\s 0 - 1\s 1)$.
What's $\partial 1 \s$? It's $11 - 10$. What's $S \partial (1 \s)$? It's
\[11\s + 1\s0 + \s00 - 10\s - 1\s1 - \s01\]
What's $(\partial S - S \partial)(1*)$? It's
\[ 11\s + 1\s 0 - 0 0 \s + 1 0 \s - 1\s 1  - 10\s    \]
What's $(\partial S - 2 \cdot S \partial)(1*)$? It's... augh, I think I messed up somewhere. I need to mechanize this.
\end{document}
