\documentclass{article}
\usepackage{amssymb}
\input{theorem}

\input{prooftree}
\def\erule#1#2{\begin{prooftree}#1\justifies #2\end{prooftree}}
\def\pair#1#2{\langle #1 , #2 \rangle}

\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{calc}
\definecolor{morange}{rgb}{1,0.56,0}
\definecolor{lorange}{rgb}{1,0.95,0.8}
\definecolor{mgreen}{rgb}{0,0.56,0}
\definecolor{lgreen}{rgb}{0.95,1,0.8}
\definecolor{mblue2}{rgb}{0,0.2,1.0}
\definecolor{lblue}{rgb}{0.8,0.95,1}
\definecolor{mred}{rgb}{0.9,0.1,0.1}
\definecolor{mgreen}{rgb}{0.1,0.5,0.1}
\definecolor{mblue}{rgb}{0.3,0.3,0.9}
\def\bitf#1{#1 [smooth, tension=0.8] coordinates {(-1.6,2) (-1,1) (0,0)}}
\def\bitg#1{#1 [smooth, tension=0.8] coordinates {(1,2) (0.55,1) (0,0)}}
\def\bitgrev#1{#1 [smooth, tension=0.8] coordinates {(0,0) (0.55,1) (1,2)}}
\def\bitfg#1{#1 [smooth, tension=0.8] coordinates {(0,0) (0,-2) }}
\def\ep{\varepsilon}
\def\bcase{\mathop\mathbf{case}}
\def\bof{\mathop\mathbf{of}}
\def\binj{\mathbf{inj}}
\def\blet{\mathrel\mathbf{let}}
\def\bin{\mathrel\mathbf{in}}
\def\bmatch{\mathrel\mathbf{match}}
\def\bwith{\mathrel\mathbf{with}}
\def\pbck{\ar[dr, phantom, pos=0, "\lrcorner"]}
\def\ups{{\uparrow}}
\def\dns{{\downarrow}}
\def\grad{\nabla}
\def\uups{{\Uparrow}}
\def\ddns{{\Downarrow}}
\def\adjust{\big|}
\def\O{\mathcal{O}}
\def\rid{\mathsf{id}}
\def\ridp{\mathsf{idp}}
\def\rcoe{\mathsf{coe}}
\def\rtype{\mathsf{type}}
%\def\int{\square}
\def\bd{\partial}
\def\prov{\vdash}
\def\prequiv{\dashv\vdash}
\def\imp{\Rightarrow}
\def\cqed{\hskip2mm{\vrule width .5em height .5em depth 0em}} % at the end of a |P.
\def\o{\circ}
\def\lx{\bigcirc}
\def\B{\mathbb{B}}
\def\C{\mathbf{C}}
\def\R{\mathbb{R}}
\def\bx{\mathbf{x}}
\def\by{\mathbf{y}}
\def\bv{\mathbf{v}}
\def\bt{\mathbf{t}}
\def\bu{\mathbf{u}}
\def\bp{\mathbf{p}}
\def\bz{\mathbf{0}}
\def\bzee{\mathbf{z}}
\def\S{\mathfrak{S}}
\def\M{\mathbb{M}}
\def\X{\mathbf{X}}
\def\Y{\mathcal{Y}}
\def\x{\times}
\def\st{\mathrel|}
\def\rset{\mathbf{Set}}
\def\rcat{\mathbf{Cat}}
\def\op{\mathsf{op}}
\def\P{\mathbb{P}}
\def\I{\mathbb{I}}
\def\U{\mathbb{U}}
\def\N{\mathbb{N}}
\def\Z{\mathbb{Z}}
\def\tw{\mathbf{2}}
\def\dash{\hbox{---}}
\def\dom{\mathop{\mathrm{dom}}}
\def\cod{\mathop{\mathrm{cod}}}
\def\celse{\mathrel{|}}
\def\cn{{:}}
\def\rok{\mathrel\mathsf{ok}}
\def\llam#1{\langle {#1} \rangle}
\def\hf{{\odot}}
\def\lerp#1#2#3{#1[\![#2,#3]\!]}
\def\nope{\emptyset}
\def\ep{\varepsilon}

\begin{document}
\tikzset{>=stealth}
\tikzset{
   commutative diagrams/.cd,
   arrow style=tikz,
   diagrams={>=stealth}}

\section{Main}
\subsection{Multiple Functions}
Let me restate an important example: the criticality behavior of $x^2 - y^2$ is properly different from the
criticality behavior of the two functions $(x-y, x+y)$. We can say that the origin is 3-critical in $f = x^2 - y^2$ because
$f =0 $ and $f_x = 0$ and $f_{xx}f_y = 0$, but more to the point it does have an extension to $\hat f = x^2 - y^2 + z$
which has, for $|z|$ arbitrarily small, multiple 2-critical points per constant-$z$ plane. Namely $(0, \pm \sqrt{\delta},  \delta)$
for small positive $\delta$. And these {\em are} definitely 2-critical points because there are multiple 1-critical values of
$x$ for a $y$ arbitrarily close to $\pm \delta$. Namely, near $(0,\sqrt{\delta}, \delta)$ there is
\[(\pm\sqrt{2\ep\sqrt{\delta} + \ep^2}, \sqrt{\delta} + \ep, \delta)\]
for arbitrarily small $\ep > 0$ and near $(0,-\sqrt{\delta}, \delta)$ there is
\[(\pm\sqrt{2\ep\sqrt{\delta} + \ep^2}, -\sqrt{\delta} - \ep, \delta)\]

However, I think the origin at the functions $(x-y,x+y)$ should be considered 3-stable. Why is this?
Well, in general suppose we have $(f,g) : \R^{2+n} \to \R^2$ such that $f = 0$ and $g = 0$ but $[f,g] = f_xg_y - f_y g_x\ne 0$. We want
to show there is a neighborhood of the origin we can restrict our attention to such that
for any $\bp \in \R^n$ there is exactly one $(x,y)$ such that $(x,y,\bp)$ is a 2-critical point.

Well, the implicit function theorem guarantees that there exists an open set $U \subseteq \R^n$ containing the origin
such that there is a unique smooth function $h : U \to \R^2$ such that
\[ h(\bz) = (0,0)\]
and
\[f(h(\bx), \bx) = 0\]
\[g(h(\bx), \bx) = 0\]
Ok, this seems like exactly what I need.

\subsection{Restating What I Know}
\begin{lemma}[Implicit Function Theorem]
If $f : \R^{m+n} \to \R^m$, and $f(\bz) = \bz$, and the jacobian of $f$ with respect to the first $m$ variables is nonsingular,
then there exists a neighborhood $U \subseteq \R^n$ of the origin, and a unique function $g : U \to \R^m$,
with $g(\bz) = \bz$ and
\[f(g(\bx), \bx) = \bz\]
for all $\bx \in U$.
\end{lemma}


Suppose I have some $p^1,\ldots,p^m : \R^n \to \R$.

If I know
\[S_1(\bx_0) := \forall i . p^i(\bx_0) \ne 0\]
 then there is a neighborhood
$U$ around $\bx_0$ such that $\forall \bx\in U. \forall i . p^i(\bx_0) \ne 0$.
This is just because of continuity. This is called 1-stability.
So 1-criticality is
\[C_1(\bx_0) := \exists i . p^i(\bx_0) = 0\]

If I have a $\bx_0 = (x_0, \by_0)$ (where $x_0\in \R$ and $\by_0 \in \R^{n-1}$) and I know that  there exists an $i$ such that

\[s_2(\bx_0) = \exists i . (\forall j\ne i . p^j(\bx_0) \ne 0) \land (p^i(\bx_0) = 0) \land (p^i_x(\bx_0) \ne 0) \]
then by the implicit function theorem for $n = n-1$ and $m=1$, there
is a neighborhood $U\subseteq \R^{n-1}$ around $\by_0$ such that for
every $\by \in U$ there exists exactly one $x \in \R$ such that
$p^i(x,\by) = 0$. Because we also assumed $(\forall j\ne i . p^j(\bx_0) \ne 0)$, this means that we can say more strongly that:
\[\exists (U\ni \by_0) \subseteq \R^{n-1}. \forall \by \in U . \exists !x \in \R . \exists i . p^i(x,\by) = 0\]
which is what we can call
\[S_2(x_0,\by_0) := \exists (U\ni \by_0) \subseteq \R^{n-1}. \forall \by \in U . \exists !x \in \R . C_1(x, \by)\]
so I can see that $s_2(\bx_0) \prov S_2(\bx_0)$.

If I take the negation of $s_2$, I get
\[c_2(\bx_0) = \forall i . (\exists j\ne i . p^j(\bx_0) = 0) \lor (p^i(\bx_0) \ne 0) \lor (p^i_x(\bx_0) = 0) \]
Now at the next dimension up, there are two distinct types of 2-critical points. One where two functions cross,
and one where a single function has a turnaround. Before I do that...

\subsection{Starting Over}
(2020.03.05)

Given a function $f : \R^n \to \R^m$, I want to compute a number $\dim f : \N$ which is
the `dimension' of the origin in the graph of $f$. We write $\hat f : \R^n \to \N$
for
\[\hat f(\bx) = \dim (\lambda \by . f(\bx + \by))\]
the dimension of the function `at' the point $\bx$.

 Let $\Pi : \R^m \to \R$ be defined by
\[\Pi(x_1, \ldots, x_m) = \prod_{i = 1}^m x_i\]
One thing I know should be true is that
\[ \dim f = 0 \qquad\iff\qquad \Pi  f(\bz) \ne 0\]


If $f : \R^{n+m} \to X$, then $f\adjust_n : \R^n \to X$ is defined by
$f\adjust_n(\bx) = f(\bx, \bz)$. We say $f \sim_n g$ if $f\adjust_n =
g\adjust_n$.

We say a predicate $P \subseteq \R^{n}\x\R^{m}$ is {\em
  $n$-locally} true if there exists neighborhoods
$U_x \subset\R^n, U_y \subset\R^m$ of $\bz$ such that
\[\forall \by\in U_y. \exists ! \bx \in U_x . P(x,y)\]

We say $\dim f = 1$ when $\dim f \ne 0$, and also for any $g : \R^{1+p} \to \R^m$ with
$g \sim_1 f$, it's 1-locally
true of any point $(x,\by)$ in $g$'s domain that $\hat g(x,\by) > 0$.

\subsubsection{Booleans Only}
Given a function $f : \R^n \to \R^m$, I want to characterize merely
whether it is `stable' at its origin or not. For $n=0$, I simply have
a value in $\R^m$. I say it's stable if no coordinate is zero.

Otherwise, I have a function $\R^{n+1} \to \R^m$. It is stable at $n+1$ if
for every extension of $f$ to $g : \R^{n+1+t} \to \R^m$, there exist neighborhoods of the origin
$ U \subseteq \R^t$ and $V \subseteq \R^{n+1}$ such that
for every $\bt \in U$, there exists a unique $(\bx,y) \in V$ such that the function
\[ \lambda \bx_0.f(\bx+\bx_0, y, \bt) : \R^n \to \R^m\]
is salient at $n$.

A function is salient at 0 if it's not stable at 0...

\subsubsection{Thinking more Categorically}
Suppose I have a category $\C$ of sets and functions. If not sets and functions, I think
I at least need finite products and zero morphisms $0: 1 \to X$. But I'm going to proceed
as if it were a concrete category. This could consist of
objects $\R^n$ and polynomial functions between them.

Suppose I have a set $C \in \C$ and a notion of which points $c\in C$ are 0-stable.
For example, $C$ might be $\R$, and all nonzero points are stable. Consider a
function $f : \R \to C$. On the one hand  $f(0)$ could be 0-stable.
Or it could be 1-like. This holds if for every extension of $f$ to
$g : \R \x D \to C$, we can find neighborhoods $U_d \subseteq D$ and $U_r \subseteq \R$
such that for all $d\in U_d$ there exists a unique $x\in U_r$ such that $g(x,d)$ is not 0-stable.
We can call this being 1-stable.

Consider a function $f : \R^2 \to C$. Any one of the following (but at most one, assuming continuity)
might be true:
\begin{enumerate}
\item $f\adjust_{x=0,y=0}$ is 0-stable.
\item $f\adjust_{y=0}$ is 1-stable.
\item For any extension of $f$ to $g: \R^2 \x D \to C$, we can find neighborhoods $U_d \subseteq D$
and $U_r\subseteq \R^2$ such that for any $d \in U_d$ there exists a unique $(x,y) \in U_r$ such
that $\lambda x_0 . f(x + x_0, y, d)$ is not 0-stable or 1-stable.
\end{enumerate}

\subsubsection{Booleans Only, Again}
Suppose $f : \R^{n+m} \to C$. We define $f\adjust_n : \R^n \to C$ by
\[f\adjust_n(\bx) = f(\bx, \bz)\]
A function $f : \R^n \to \R^m$ may or may not be {\em stable}, defined inductively on $n$.

For $n=0$, the function $f$ is simply a value in $\R^m$. We say it's stable if no coordinate is zero.

Otherwise, I have a function $\R^{n+1} \to \R^m$. It considered stable if
\begin{tabbing}
For any $u \in \N$\\
For any polynomial $g : \R^{n+1+u} \to \R^m$ such that $g\adjust_{n+1} = f$\\
There exists a neighborhood $U \subseteq \R^u$ of $\bz$ and\\
There exists a neighborhood $V \subseteq \R^{n+1}$ of $\bz$ such that\\
For any $\bu \in U$\\
There exists a unique $(\bx_0,y_0)\in V$ such that\\
For any $k \in [0,n]$\\
The function $(\lambda \bx.g(\bx+\bx_0, y_0, \bu))\adjust_k : \R^k \to \R^m$ is not stable.
\end{tabbing}

\subsubsection{Generically, Again}

We have some kind of set $C$ and a notion of which points $c\in C$ are stable,
and which functions $\R^n\to C$ are `allowed'.

We aim to define which functions $f : \R^n \to C$ are stable.
For $n=0$, the function $f$ is simply a value in $C$, so we inherit the provided definition.

Otherwise, we have a function $\R^{n+1} \to C$. It considered stable if
\begin{tabbing}
For any $u \in \N$\\
For any allowable path $g : \R^{n+1+u} \to C$ such that $g\adjust_{n+1} = f$\\
There exists a neighborhood $U \subseteq \R^u$ of $\bz$ and\\
There exists a neighborhood $V \subseteq \R^{n+1}$ of $\bz$ such that\\
For any $\bu \in U$\\
There exists a unique $(\bx_0,y_0)\in V$ such that\\
For any $k \in [0,n]$\\
The function $(\lambda \bx.g(\bx+\bx_0, y_0, \bu))\adjust_k : \R^k \to C$ is not stable.
\end{tabbing}

\subsubsection{With Dimensions, Again}

A {\em bundle} is a
\begin{enumerate}
\item A set $C$
\item A predicate $\chi \subseteq C $
\item A notion $\alpha \subseteq \R^n \to C$ of which functions $\R^n \to C$ are `allowed'.
\end{enumerate}

We want to lift one bundle $(C, \chi, \alpha)$ to another one $(C', \chi', \alpha')$
 whose carrier is $C' = \R \to C$.

Given $f : \R \to C$, we say $f$ is critical $(\in \chi')$ if $f(0) \in \chi$ and not
\begin{tabbing}
For any $u \in \N$\\
For any allowable function $g : \R \x \R^u \to C$ such that $g\adjust_{1} = f$\\
There exists a neighborhood $U \subseteq \R^u$ of $\bz$ and\\
There exists a neighborhood $V \subseteq \R$ of $\bz$ such that\\
For any $\bu \in U$\\
There exists a unique $v \in V$ such that\\
$g(v, \bu) \in \chi$.
\end{tabbing}

Start with $C = \R$ and $\chi = \{0\}$. What is $(C,\chi)'$?
The set $C'$ is $\R \to \R$. What is $\chi'$? We require of any $f \in \chi'$ that $f(0) = 0$.
But also that it's not the case that any extension has the property that there's locally
one zero.

This doesn't seem right, though; I'm not quantifying unique existence over a growing $\R^n$.

\subsubsection{Capturing the Domain Correctly?}
Forget for the moment about $\alpha$.

A {\em bundle} is a
\begin{enumerate}
\item A vector space $D$
\item A set $C$
\item A predicate $\chi \subseteq D \to C $ that says which diagrams are `critical'.
\end{enumerate}
What's the induction step? We pass from $(D, C, \chi)$ to $(D \x \R, C, \chi')$.
Some $f : D \x \R \to C$ is going to be in $\chi'$ if, first of all, $f(\dash, 0) \in \chi$,
and secondly, it's {\em not} the case that it's stable --- where $f$ is stable iff
\begin{tabbing}
For any appropriate vector space $\mathfrak U$\\
For any allowable function $g : D \x \R \x {\mathfrak U} \to C$ such that $g(\dash,\dash,\bz) = f$\\
There exists a neighborhood $U \subseteq {\mathfrak U}$ of $\bz$ and\\
There exists a neighborhood $V \subseteq D\x \R$ of $\bz$ such that\\
For any $\bu \in U$\\
There exists a unique $(d, v) \in V$ such that\\
$g(d + \dash, v, \bu) \in \chi$.
\end{tabbing}

\section{A Digression}

bblum mentioned the idea of bisecting a commit history to find the commit that first introduces
a flaky test failure. The continuous version of the problem seems easier to manage. Suppose
we have a random variable $\phi$ that lives the interval $[0,1]$. This is where the test
first fails. If we probe the system at some $t \ge \phi$, then there is a probability $p$
of detecting a failure. If we probe the system at $t < \phi$, then there is probability 0
of failure.
\[ P[F_t \st \phi=\phi ] = \cases{p & $t \ge \phi$  \cr 0 & $ t < \phi$ }\]
So we're interested in the Bayesian update when we pick a $t$ to probe at, and see a failure
or not. Suppose we have a prior distribution $f(\phi)$ on $\phi$.
\[ P[\phi = \phi \st F_t] = {P[ F_t \st \phi=\phi ]  \over P[F_t]} P[\phi=\phi]\]
\[  = {P[ F_t \st \phi=\phi ]  \over P[F_t]} f(\phi)\]
\[  = {P[ F_t \st \phi=\phi ]f(\phi)  \over {\int_0^1 P[ F_t \st \phi=\phi ]f(\phi) d\phi  } } \]
Let that normalizing constant be $K = \int_0^1 P[ F_t \st \phi=\phi ]f(\phi) d\phi$. Then
\[ P[\phi = \phi \st F_t] =
\cases{
pf(\phi)/K & $\phi \le t$  \cr 0 & $ \phi > t$
}
\]
So in this case the $p$ gets absorbed into the normalizing constant; it's just
\[ P[\phi = \phi \st F_t] =
\cases{
f(\phi)/\int_0^t f(\phi) & $\phi \le t$  \cr 0 & $ \phi > t$
}
\]
Ok, great, what about when the test seems to pass? $P[\phi = \phi | \lnot F_t]$?
That's going to be
\[   {P[ \lnot F_t \st \phi=\phi ]f(\phi)  \over K } \]
so
\[ P[\phi = \phi \st \lnot F_t] =
\cases{
(1-p)f(\phi)/K & $\phi \le t$  \cr f(\phi)/K & $ \phi > t$
}
\]
and here
\[K = (1-p)\int_0^{t} f(\phi) d\phi + \int_t^1 f(\phi) d\phi\]
\[K = 1 -p\int_0^{t} f(\phi) d\phi \]

\subsection{Entropy Calculations}
Let's say I have a nonnormalized distribution $f(\phi)$. So the real distribution
is $F(\phi) = f(\phi) /  \int_0^1 f(\phi)d\phi$. What is its entropy?
Let's suppose we can get away with differential entropy
\[ - \int_0^1 F(\phi) \ln F(\phi) d\phi \]
\[ =  {- \int_0^1 f(\phi) \ln F(\phi) d\phi \over \int_0^1 f(\phi)d\phi}\]
\[ =  {- \int_0^1 f(\phi) \ln \left( f(\phi) /  \int_0^1 f(\phi)d\phi \right) d\phi \over \int_0^1 f(\phi)d\phi}\]
\[ =  {- \int_0^1 f(\phi) \left( \ln ( f(\phi)) - \ln ( \int_0^1 f(\phi)d\phi) \right) d\phi \over \int_0^1 f(\phi)d\phi}\]
\[ =  {
- \int_0^1 f(\phi)  \ln ( f(\phi))   d\phi
+ \int_0^1 f(\phi)   \ln ( \int_0^1 f(\phi)d\phi)  d\phi
 \over \int_0^1 f(\phi)d\phi}\]
\[ =  {
- \int_0^1 f(\phi)  \ln ( f(\phi))   d\phi
 \over \int_0^1 f(\phi)d\phi}
+  \ln \left( \int_0^1 f(\phi)d\phi\right)  \]
Ok, so if we say $Z_f =  \int_0^1 f(\phi)d\phi$ and $H_F$ is the entropy of $F$, then
\[ H_F = H_f / Z_f + \ln (Z_f)\]

\subsection{Trying to work out the first choice}

Let's start with the uniform distribution. If I test at $t$, the probability of being after
the bad commit is $t$, and so the probability of seeing a failure is $pt$. So with probability
$pt$, I get the distribution
\[ P[\phi = \phi \st F_t] =
\cases{
f(\phi)/\int_0^t f(\phi) & $\phi \le t$  \cr 0 & $ \phi > t$
}
\]
\[ =
\cases{
1/t & $\phi \le t$  \cr 0 & $ \phi > t$
}
\]
which has entropy
\[ -\int_0^t 1/t \ln(1/t) d\phi = \ln(t)\]
and with probability $1-pt$ I get the distribution
\[ P[\phi = \phi \st \lnot F_t] =
\cases{
(1-p)f(\phi)/K & $\phi \le t$  \cr f(\phi)/K & $ \phi > t$
}
\]
\[  =
\cases{
(1-p)/K & $\phi \le t$  \cr 1/K & $ \phi > t$
}
\]
\[  =
\cases{
(1-p)/(1-pt) & $\phi \le t$  \cr 1/(1-pt) & $ \phi > t$
}
\]
which has entropy
\[ -\int_0^t {1-p \over 1-pt} \ln\left({1-p \over 1-pt}\right) d\phi
 -\int_t^1 {1 \over 1-pt} \ln\left({1 \over 1-pt}\right) d\phi \]
\[ =  - {t-pt \over 1-pt} \ln\left({1-p \over 1-pt}\right)
- {1-t \over 1-pt} \ln\left({1 \over 1-pt}\right)  \]
\[ =   {t-pt \over 1-pt} \ln\left({ 1-pt\over 1 -p }\right)
+ {1-t \over 1-pt} \ln\left( 1-pt \right)  \]
\[ = \left(1\over 1 - pt\right)\left(  (t-pt) \ln\left({ 1-pt\over 1 -p }\right)
+ (1-t) \ln\left( 1-pt \right) \right) \]
\[ = \left(1\over 1 - pt\right)\left(
 (t-pt) \ln\left({ 1-pt }\right)
- (t-pt) \ln\left({ 1 -p }\right)
+ (1-t) \ln\left( 1-pt \right) \right) \]
\[ = \left(1\over 1 - pt\right)\left(
- (t-pt) \ln\left({ 1 -p }\right)
+ (1-pt) \ln\left( 1-pt \right) \right) \]
\[ =
- t{1-p \over 1 - pt} \ln\left({ 1 -p }\right)
+  \ln\left( 1-pt \right)  \]

So my expected entropy overall is

\[ H = pt \ln(t) + (1-pt) \ln (1 - pt) - t(1-p) \ln (t-p) \]
and if I want to minimize this, I want to set the $t$ derivative to zero, so
\[ H_t = p \ln(t) -p \ln (1 - pt) - (1-p) \ln (t-p) + \]
\[ p + {{1-pt \over -p} \over 1 - pt} - {t(1-p) \over t-p} \]
which looks like a giant mess that does not admit an analytic solution for $t$.

\section{Back to Analytic Category Theory}

Here's a weak definition of stability for a point in the domain of a function $f : \R^n \to \R^p$.

\begin{itemize}
\item
  say $\bx \in \R^n$ is 0-stable if $f(\bx)$ has no 0-coordinates.
\item If $q > 0$, say a point is $q$-stable if there is a neighborhood $X \x Y \subseteq \R^q \x \R^{n-q}$
around it such that for every $\by \in Y$ there exists at most one $\bx \in X$ such that $(\bx,\by)$ isn't
$(q-1)$-stable.
\end{itemize}

I think it's a lemma that if you're $q$-stable, you're $(q+1)$-stable,
because every $q$-stable point has a neighborhood that is entirely
$q$-stable.

Can I compact the statement of the implicit function theorem fruitfully?

\subsection{Directed Space}

I want to say a directed space is a choice of $p \in \N$ together with, for every $n \in \N$, a semialgebraic set over the
formal derivatives and projections of functions $f : \R^n \to \R^p$.

For example, I might want to impose, for $n = 2$, that $\lnot(f^1 = 0 \lor f^2 = 0)$.

This is going to get infinite pretty fast. How can I make it more finite? If I want to deal only in 1-categorical behavior,
I can limit to first derivatives. Consider $\R^3$ for example, and the constraints I can make are logical formulae involving
$f : \R \to \R^3$ and $f_x : \R \to \R^3$. I could require that $f^1 = 0 \imp f^1_x > 0$.


\end{document}
