\documentclass{article}
\usepackage{setspace}

\usepackage[framestyle=fbox,framefit=yes,heightadjust=all,framearound=all]{floatrow}
\usepackage{amssymb}
\input theorem
\input prooftree
\def\nto{\mathrel{\dot\to}}
\def\sem#1{[\![#1]\!]}
\def\seme#1{[\![#1]\!]_{\eta}}
\def\mode{\models_{M,\eta}}
\def\C{\mathbf{C}}
\def\D{\mathbf{D}}
\def\P{\mathbf{P}}
\def\N{\mathbf{N}}
\def\tri{\triangleright}
\def\x{\times}
\def\rset{\mathbf{Set}}
\def\gap{\hskip 1em\hfil\hskip 1em}
\def\sgap{\hskip 0.3em\hfil\hskip 0.3em}
\def\XXX#1{{\color{red}\em XXX #1}}
\def\celse{\mathrel|}
\def\st{|}
\def\cqed{\hskip2mm{\vrule width .5em height .5em depth 0em}} % at the end of a |P.
\def\imp{\Rightarrow}
\def\tensor{\otimes}
\def\htensor{\mathrel{\hat\otimes}}
\def\lol{\multimap}
\def\hlol{\mathrel{\hat\multimap}}
\def\op{\mathrm{op}}
\def\prov{\vdash}
\def\prequiv{\dashv\vdash}
\def\susp#1{\langle {#1} \rangle}
\def\erule#1#2{
\begin{prooftree}
#1 \justifies #2
\end{prooftree}}
\def\cn{{:}}
\def\ups{{\uparrow}}
\def\dns{{\downarrow}}
\def\amp{\mathrel{\&}}
\def\dash{\hbox{---}}


\title{Categorical Semantics of Focused ILL}
\author{Jason Reed}
\bibliographystyle{alpha}
\begin{document}

\maketitle
\section{Introduction}
I want to describe a certain categorical semantics of focused
intuitionistic linear logic so that I can better understand how it
relates to other semantics I've seen in the literature \cite{polarized.mall}.

\section{Language}
We recall the syntax of focused intuitionistic linear logic.
Propositions are polarized into positive and negative.
There are shift operators $\ups$ and $\dns$
that coerce back and forth between the two polarities.
Atomic propositions also come in positive $a^+$ and $a^-$. Somewhat unusually (but it's just a matter of presentation,
not an essential part of the result)
we distinguish an atomic proposition $a^\pm$ of either polarity from
the suspension $\susp {a^\pm}$ of it that arises after asynchronous decomposition terminates at it.
$$\begin{tabular}{rlll}
Positives&$P$&$::=$&$ \dns N \celse P \tensor P \celse P \oplus P \celse 1 \celse 0 \celse a^+ $\\
Negatives&$N$&$::=$&$\ups P \celse  P \lol N \celse N \amp N \celse \top  \celse a^-$\\
Positive Contexts&$\Omega$&$::=$&$\cdot \celse P, \Omega$\\
Negative Contexts&$\Gamma$&$::=$&$\cdot \celse \Gamma, H$\\
Stable Hypotheses&$H$&$::=$&$N \celse \susp{a^+}$\\
Stable Conclusions&$Q$&$::=$&$P \celse \susp{a^-}$\\
Conclusions&$R$&$::=$&$N \celse Q$\\
\end{tabular}$$

The three judgments of the logic are
\[\begin{tabular}{rl}
Inversion&$\Gamma;\Omega\prov R$\\
Right Focus&$\Gamma\prov [P]$\\
Left Focus&$\Gamma[N]\prov Q$\\
\end{tabular}\]
(we sometimes abbreviate $\Gamma;\cdot \prov R$ as $\Gamma\prov R$)
and the proof rules for the focusing system are in
Figure~\ref{foc-rules}.
\input{foc-rules}

\subsection{Results}
These are thoroughly standard.

\begin{lemma}[Cut]\
  \begin{enumerate}
  \item If $\Gamma_1[N] \prov Q$ and $\Gamma_2 \prov N$, then
    $\Gamma_1, \Gamma_2 \prov Q$.
  \item If $\Gamma_1 \prov [P]$ and $\Gamma_2; P \prov Q$, then
  $\Gamma_1, \Gamma_2 \prov Q$.
  \end{enumerate}
\end{lemma}

\begin{lemma}[Identity]\
  \begin{enumerate}
  \item $N \prov N$
  \item $P \prov P$
  \item If $\Gamma_0[N] \prov Q$ implies
    $\Gamma_0, \Gamma_\bullet \prov Q$
    for every $\Gamma_0$ and $Q$, then  $\Gamma_\bullet \prov N$.
  \item If $\Gamma_0 \prov [P]$ implies $\Gamma_0, \Gamma_\bullet; \Omega \prov R$
    for every $\Gamma_0$, then $\Gamma_\bullet; P, \Omega \prov R$.
  \end{enumerate}

\end{lemma}

\section{Semantics}

A {\em model}  $M$ consists of the following data:
\begin{enumerate}
\item a symmetric monoidal category $\P$ and a category $\N$
\item functors $\_\lol\_ : \P \x \N \to \N$ and $\_\tri\_ : \P \x \N \to \rset$
\item a natural isomorphism
$$
k :
\begin{prooftree}
(P_1 \tensor P_2) \tri N
\Justifies
P_1  \tri (P_2 \lol N)
\end{prooftree}
$$
\end{enumerate}

We write $\hat \C$ as an abbreviation for $\C \to \rset$, despite the
usual contravariant convention. Given an environment, i.e. a mapping
$\eta$ of atoms $a^+$ (resp. $a^-$) to objects of $\hat \P$ (resp.
$\hat \N$), we inductively define an interpretation of all positive
(resp. negative) propositions $P$ (resp. $N$)
 as objects $\seme{P} \in \hat \P$ (resp. $\seme{N} \in \hat \N$).
We write $\sem\dash$ instead of $\seme\dash$ when the $\eta$ is evident from context.
The definition of the interpretation is given in Figure~\ref{semantics},
where $\cup$ denotes the coproduct and $\emptyset$ the initial object,
 in the category $\hat P$ or $\hat N$ as appropriate. The arrow $\to$ in the definition
of the shifts is just the  function space in $\rset$.
The multiplicative connectives are defined with coends, using the
operators $\hlol$ and $\htensor$ defined in Figure~\ref{semops}.
Shifts are defined with ends.
\begin{figure}
\begin{center}
\[
\erule{X \in \hat\P \qquad Y \in \hat\N}
{X \hlol Y := \int^{\alpha : \P, \phi: \N} X(\alpha) \x Y(\phi) \x \N(\alpha \lol \phi, \dash)}
\]
\[
\erule{X \in \hat\P \qquad Y \in \hat\P}
{X \htensor Y := \int^{\alpha_1 : \P, \alpha_2: \P} X(\alpha_1) \x Y(\alpha_2) \x \P(\alpha_1 \tensor \alpha_2, \dash)}
\]
\end{center}
\caption{Semantic Operations \label{semops}}
\end{figure}

\begin{figure}
\begin{center}
\[\sem{a^\pm} = \eta(a^\pm) \qquad \qquad \sem{1} = \P(1, \dash)\]
\[\sem{P_1 \oplus P_2} =  \sem{P_1} \cup \sem{P_2}\qquad
\sem{0} = \emptyset\]
\[\sem{N_1 \amp N_2} =  \sem{N_1} \cup \sem{N_2}\qquad
\sem{\top} = \emptyset\]
\[\sem{P \lol N} = \sem P \hlol \sem N\]
\[\sem{P_1 \tensor P_2} = \sem {P_1} \htensor \sem{P_2}\]
\[\sem{\ups P } = \int_{\alpha : \P} \sem P (\alpha) \to (\alpha \tri \dash)\]
\[\sem{\dns N } = \int_{\phi : \N} \sem N (\phi) \to (\dash \tri \phi)\]
\[\erule{\Gamma = H_1, \ldots, H_n}{\sem{\Gamma} = \sem{ H_1}^< \htensor \cdots \htensor \sem{ H_n}^<}
\qquad
\erule{\Omega = P_1, \ldots, P_n}{\sem{\Omega} = \sem{P_1}^< \htensor \cdots \htensor \sem{P_n}^<}\]
\[\sem P^< = \sem{P} \qquad \sem N^< = \sem{\dns N} \qquad \sem{\susp{a^+}}^< = \eta(a^+)\]
\[\sem N^> = \sem{N} \qquad \sem P^> = \sem{\ups P} \qquad \sem{\susp{a^-}}^> = \eta(a^-)\]
\end{center}
\caption{Interpreting Propositions \label{semantics}}
\end{figure}


To interpret sequents, do as follows. If $X : \hat \C$ and $Y : \hat \D$, define their
objectwise product
$X \land Y : \widehat{\C \x \D}$ as $(X \land Y) (C, D) =  X(C) \x Y(D)$.

We say
\begin{enumerate}
\item  $\Gamma;\Omega \mode R$ iff there is a nat. trans. $(\sem \Gamma \htensor \sem \Omega) \land \sem{R}^> \nto \tri$
\item  $\Gamma \mode [P]$ iff there is a nat. trans. $\sem \Gamma   \nto \sem{ P}$
\item  $\Gamma[N] \mode Q$ iff there is a nat. trans. $(\sem \Gamma \hlol \sem Q^>) \nto \sem N$
\end{enumerate}
When we drop the subscripts $M, \eta$ and just write $\models$, it  means that the statement
holds for all $M, \eta$.


\section{Results}

\begin{theorem}[Soundness]\
  \begin{enumerate}
  \item If $\Gamma; \Omega \prov R$ then $\Gamma ; \Omega \models R$
  \item If $\Gamma \prov [P]$ then $\Gamma \models [P]$
  \item If $\Gamma[N] \prov Q$ then $\Gamma[N] \models Q$
  \end{enumerate}
\end{theorem}

\begin{proof}
  By induction on the derivation.
\cqed
\end{proof}




\begin{theorem}[Completeness]\
  \begin{enumerate}
  \item If $\Gamma; \Omega \models R$ then $\Gamma ; \Omega \prov R$
  \item If $\Gamma \models [P]$ then $\Gamma \prov [P]$
  \item If $\Gamma[N] \models Q$ then $\Gamma[N] \prov Q$
  \end{enumerate}
\end{theorem}

\begin{proof}
  We build a universal model. Fix a countable collection of atoms.
  \begin{itemize}
  \item Let $\P$ be the category whose objects are all contexts $\Gamma$, and
whose morphisms are permutations.
The functor $\tensor$ is multiset union, and the unit $1$ is the empty context.
\item Let $\N$ be the category whose objects are pairs $(\Gamma, Q)$ and whose
morphisms are context permutations.
\item
The functor $\lol$ takes an object $\Gamma_1$ of $\P$ and an object $(\Gamma_2, Q)$ of $\N$
and produces $((\Gamma_1, \Gamma_2), Q)$ in $\N$.
\item The functor $\tri$ takes an object $\Gamma_1$ of $\P$ and an object $(\Gamma_2, Q)$ of $\N$
and yields the set of derivations $\Gamma_1, \Gamma_2 \prov Q$.
\item
  We choose as interpretations of the atoms as follows
\[\eta(a^+)(X: \P) = \cases{\{*\}&if $X = (\susp{a^+})$; \cr \{\} & otherwise.}\]
\[\eta(a^-)(X: \N) = \cases{\{*\}&if $X = (\cdot, \susp{a^-})$; \cr \{\} & otherwise.}\]
  \end{itemize}
  The thing we need to do is show that this model is in fact universal ---
  that the interpretation of every proposition reflects its (focused)
  provability. We claim that for all $P$ and $N$ we have
  \begin{itemize}
  \item $\sem{P}(\Gamma) \iff  \Gamma \prov [P]$
  \item $\sem{N}(\Gamma, Q) \iff \Gamma[N] \prov Q$
  \end{itemize}

\cqed
\end{proof}

\bibliography{../wes}

\end{document}
