\documentclass{article}
\input{theorem}
\input{prooftree}
\usepackage{stmaryrd}
\usepackage{latexsym}
\usepackage{yfonts}
\usepackage{amsmath}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{tikz}
\usetikzlibrary{calc,arrows,cd,decorations.pathreplacing}
\usetikzlibrary{decorations.pathmorphing}
\usepackage{tcolorbox}
\tcbuselibrary{breakable}
\definecolor{cyan}{rgb}{0.80,0.95,0.95}
\definecolor{orange}{rgb}{0.95,0.90,0.80}
\definecolor{lred}{rgb}{0.95,0.8,0.8}
\definecolor{lgray}{gray}{0.7}
\newtcolorbox{calldefn}{colframe=lred,colback=lred,grow to right by=-10mm,grow to left by=-10mm,
boxrule=0pt,boxsep=0pt,breakable}
\newtcolorbox{callout}{colframe=cyan,colback=cyan,grow to right by=-10mm,grow to left by=-10mm,
boxrule=0pt,boxsep=0pt,breakable}
\newtcolorbox{callout2}{colframe=orange,colback=orange,grow to right by=-10mm,grow to left by=-10mm,
boxrule=0pt,boxsep=0pt,breakable}

\input{linear}
\def\rfst{\mathsf{fst}\,}
\def\rsnd{\mathsf{snd}\,}
\def\bunfold{\mathbf{unfold}}
\def\bfold{\mathbf{fold}}
\def\susp#1{\langle {#1} \rangle}
\def\ssusp#1{\langle\!\langle {#1} \rangle\!\rangle}
\def\dto{\mathrel{\dot\to}}
\def\dns{{\downarrow}}
\def\ups{{\uparrow}}
\def\sprov{\Vdash}
\def\signat{{\color{red}\Sigma}}

\def\wat{\mathbin{@}}
\def\wwat{\mathbin{\underline\wat}}
\def\cn{{:}}
\def\dv{{\div}}
\def\ww{{\mathsf w}}
\def\base{{\textfrak b}}

\def\N{{\mathbb N}}
\def\munit{I}
\def\tt{{\mathsf t}}
\def\rr{{\mathsf r}}
\def\ii{{\mathbb I}}
\def\pp{\textsf{\textbf p}}
\def\nn{{\mathsf n}}
\def\PP{\textsf{\textbf P}}
\def\NN{{\mathsf N}}
\def\zz{{\mathsf z}}
\def\sem#1{[\![#1]\!]}
\def\usem#1{\lceil\!\!\lceil{#1}\rceil\!\!\rceil}
\def\zem#1{\langle\!\langle#1\rangle\!\rangle}
\def\col#1{{\mathsf{C}}_{#1}}
\def\lax{\mathop\bigcirc}
\def\rjust{\mathsf{just}}
\def\rcvt{\mathsf{cvt}}
\def\rtype{\mathsf{type}}
\def\rkind{\mathsf{kind}}
\def\rmode{\mathsf{mode}}
\def\rprel{\mathbf{Prel}}
\def\rrfam{\mathbf{Rfam}}
\def\rset{\mathbf{Set}}
\def\rrel{\mathbf{Rel}}
\def\rcat{\mathbf{Cat}}
\def\rfam{\mathbf{Fam}}
\def\binj{\mathbf{inj}}
\def\rid{\mathsf{id}}
\def\rtrue{\mathrel\mathrm{true}}
\def\pbck{\ar[dr, phantom, pos=0, "\lrcorner"]}
\def\bpbck{\ar[ddrr, phantom, pos=0, "\lrcorner"]}
\def\erule#1#2{\begin{prooftree}#1\justifies #2\end{prooftree}}
\def\lpar{\bindnasrepma}
\def\lamp{\binampersand}
\def\btwo{\mathbf{2}}
\def\bone{\mathbf{1}}

\newarrow {Equiv} =====
\def\FinSet{\mathbf{FinSet}}
\def\x{\times}
\def\o{\comp}
\def\maps{\ar[dd,mapsto,shorten <=2em, shorten >=2em]}
\def\mapt{\ar[ddd,mapsto,shorten <=2em, shorten >=2em]}
\def\frommaps{\ar[from=uu,mapsto,shorten <=2em, shorten >=2em, crossing over]}
\def\frommapt{\ar[from=uuu,mapsto,shorten <=2em, shorten >=2em, crossing over]}
\def\C{\mathbf{C}}
\def\S{\mathbf{S}}
\def\D{\mathbf{D}}
\def\E{\mathbf{E}}
\def\M{\mathbf{M}}
\def\pair#1#2{\langle#1,#2\rangle}
\def\tri{\triangleright}
\def\o{\circ}
\def\rctx{\,\mathsf{ctx}}
\def\del{\partial}
\def\also#1{\ \textcolor{blue}{\celse #1}}
\newcounter{nodemaker}
\setcounter{nodemaker}{0}
\def\twocell#1#2{%
  \global\edef\mynodeone{twocell\arabic{nodemaker}}%
  \stepcounter{nodemaker}%
  \global\edef\mynodetwo{twocell\arabic{nodemaker}}%
  \stepcounter{nodemaker}%
  \ar[#1,phantom,shift left=3,""{name=\mynodeone}]%
  \ar[#1,phantom,shift right=3,""'{name=\mynodetwo}]%
  \ar[Rightarrow,#2,from=\mynodeone,to=\mynodetwo]%
}
\def\twocellswap#1#2{%
  \global\edef\mynodeone{twocell\arabic{nodemaker}}%
  \stepcounter{nodemaker}%
  \global\edef\mynodetwo{twocell\arabic{nodemaker}}%
  \stepcounter{nodemaker}%
  \ar[#1,phantom,shift right=3,""{name=\mynodeone}]%
  \ar[#1,phantom,shift left=3,""'{name=\mynodetwo}]%
  \ar[Rightarrow,#2,from=\mynodeone,to=\mynodetwo]%
}
\def\null{\emptyset}
\def\maps{\ar[dd,mapsto,shorten <=2em, shorten >=2em]}
\author{Jason Reed}
\definecolor{fib}{HTML}{ff7f00}
\definecolor{green}{HTML}{007f00}
\definecolor{red}{HTML}{df3f3f}
\definecolor{opfib}{HTML}{007fff}

% For lightening colors I used
% http://trycolors.com/ 6 parts white to 1 part color
\definecolor{opfibl}{HTML}{DBEDFF}
\definecolor{greenl}{HTML}{DBEDDB}
\definecolor{redl}{HTML}{FAE4E4}
\def\njudge#1{\fcolorbox{opfib}{white}{#1}}
\def\nrule#1{\fcolorbox{white}{opfibl}{#1}}
\def\neqn#1{\fcolorbox{white}{redl}{#1}}



\def\judge#1{\vskip 2em\noindent \njudge{$#1$}\vskip 0.5em\noindent}
\def\rule#1{\vskip 2em \noindent \nrule{$#1$}\vskip 0.5em\noindent}
\def\eqn#1{\vskip 2em \noindent \neqn{$#1$}\vskip 0.5em\noindent}
\def\prof{\nrightarrow}
\def\tensor{\otimes}
\def\htensor{\mathrel{\hat\otimes}}
\def\lol{\multimap}
\def\hlol{\mathrel{\hat\multimap}}
\def\rpath{\mathsf{Path}}
\def\wtf{{\color{red}???}}
\begin{document}

\def\
\tikzset{
   commutative diagrams/.cd,
   arrow style=tikz,
   diagrams={>=stealth}}
% got this from
% https://tex.stackexchange.com/questions/169512/tikz-style-arrow-tips-missing-when-using-tikz-cd-crossing-over
% ???

\section{Some Notes on What I Am Trying to Do}

There is a difference in flavor I perceive between two sorts of
type-theory work. They may not be fundamentally distinct --- I would
be only too happy if someone could explain to me how they could be
unified or better connected or whatever --- but for the time being, I
feel like giving them names, to at least explain my feelings about them.

One is the practice of {\em external type theory}: discovering new
ways to build and combine types, learning about how types interact
with one another. I put in this category the study of linear type
systems, modal type systems, and really most of the ordinary,
essential basics of type theory as we use them day-to-day in
programming as well, algebraic data types, existential types,
universal types. A canonical achievement of external type theory is to
describe a new logical connective, a new type-former.

The other is {\em internal type theory}, which is instead about
studying what is (or what can be) going on {\em internally} to a type.
The phenomenon of parametricity is a phenomenon of internal type
theory, because it supposes that types come {\em equipped} with
relations. `Distance makes the types grow stronger' is a work of
internal type theory because it supposes that types come equipped with
metrics. Intuitionism per se feels like an internal perspective
because it tells us that types are not merely sets, but sets evolving
through time. Homotopy type theory is an internal-flavored approach,
because it tells us that types are not merely types but {\em spaces}.
A canonical achievement of internal type theory is to describe a
richer category in which programming can be done.
\end{document}
